---
title: "Endowment Time Series"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    toc: true
    toc_depth: 4
    toc_float: true 
    css: !expr here::here('css', 'template.css')
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "./output_html")})
---

<style type="text/css">
.main-container {
  max-width: 2200px !important;
  margin-left: auto;
  margin-right: auto;
}
</style>


This file contains interactive `plotly` plots on endowment data over time.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE, comment=NA, eval=FALSE)
library(tidyverse)
library(kableExtra)
library(scales)
library(plotly)
library(glue)
library(here)
library(viridis)


all_plots <- FALSE

m <- list(
    l = 50,
    r = 50,
    b = 200,
    t = 150,
    pad = 0.5
)


theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change
    
    theme(
      
      
      #text elements
      plot.title = element_text(             #title
                   size = 14,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),               
      
      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size
      
      axis.title = element_text(             #axis titles
                   size =14),               #font size
      
      axis.text.x = element_text(              #axis text
                   size = 12),
      legend.text = element_text(size = 10),
      legend.title = element_text(size = 11, face="bold"),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 14, face = "bold", color="white"),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)
   
}




```

```{r load data, eval=TRUE}

# load companies file of EIN to name and endowment data

companies_to_ein <- read_rds(here('data',   'companies.RDS'))

companies_to_ein <- companies_to_ein %>%
  mutate(organization_name = ifelse(
    organization_name == "Ballet Hispanico", 
    "Ballet Hispánico",
    organization_name))

endowment_data <- read_rds(here("data", 
                                "endowments_by_most_recent_filings.RDS")) %>%
  select(-c(EndowmentsHeldUnrelatedOrgInd, EndowmentsHeldRelatedOrgInd)) %>%
  pivot_longer(-c(EIN, fiscal_year),
               names_to = "variable_name") %>%
  left_join(companies_to_ein) %>%
  mutate(fiscal_year=as.numeric(paste(fiscal_year)))

```

```{r,eval=FALSE}

# extract return dates
source(here("GET_VARS.R"))

files <- dir(here("ballet_990_released_20230208"),
              full.names = TRUE)


dates <- map_df(files,
                ~get_df(filename = .x, 
                        variables = c("//Return//ReturnHeader//TaxPeriodEndDt"))) %>%
  mutate(fiscal_year = as.numeric(paste(fiscal_year))) %>%
   filter_ein()

saveRDS(dates, here('data', 'dates.RDS')) 


```

```{r, eval=TRUE}

dates <- readRDS( here('data', 'dates.RDS')) %>%
  select(EIN, TaxPeriodEndDt, fiscal_year) 
 

endowment_data <- endowment_data %>%
  mutate(fiscal_year=as.numeric(paste(fiscal_year))) %>%
  left_join(dates)


```


```{r, eval=TRUE}

# function to plot variables of interest against each other
plot_ranks <- function(var1, var2, data) {

  
   plt <- data %>%
    group_by(fiscal_year) %>%
   # arrange(var1) %>%
    mutate("{var1}_rank" := rank(-!!sym(var1)), na.last = "keep") %>%
#    arrange(var2) %>%
    mutate("{var2}_rank"  := rank(-!!sym(var2)),  na.last = "keep") %>%
    ggplot(aes(x = !!sym(glue("{var1}_rank" )), y =!!sym(glue("{var2}_rank" )),
               color  = organization_name,
               label =EIN
               )) +
    geom_point() +
    geom_function(fun=function(x)x,color="darkred", alpha = .8) +
    labs(x = paste0(var1, " Rank"),
         y =  paste0(var2, " Rank")) +
    theme_bw() +
    labs(title = glue("Rank of {var2} vs. Rank of {var1}")) +
    viridis::scale_color_viridis(discrete=TRUE,
                                 option = "rocket",
                                 end = .9) +
     facet_wrap(~fiscal_year)+
      theme(plot.title = element_text(size = 14, 
                                      hjust = .5, face="bold",),
            plot.subtitle = element_text(hjust = .5, 
                                         face="italic",
                                         size = 14),
            axis.title = element_text(size = 13, 
                                      face = "bold")) 
  
  ggplotly(plt, margin = m, height = 550) %>%
    partial_bundle()

}

plot_ranks_by_consistency <- function(var1, var2, data) {

  
   plt <- data %>%
     filter(fiscal_year > 2010 & fiscal_year < 2021) %>%
    group_by(fiscal_year) %>%
   # arrange(var1) %>%
    mutate("{var1}_rank" := rank(-!!sym(var1)), na.last = "keep") %>%
#    arrange(var2) %>%
    mutate("{var2}_rank"  := rank(-!!sym(var2)),  na.last = "keep")  %>%
     mutate(rank_diff = !!sym(glue("{var2}_rank")) - !!sym(glue("{var1}_rank" ))) %>%
   group_by(EIN) %>%
   mutate(sum_pos = sum(rank_diff >0, na.rm=TRUE),
          sum_neg = sum(rank_diff < 0,  na.rm=TRUE),
          sum_zero = sum(rank_diff ==0, na.rm=TRUE))%>%
     ungroup() %>% 
   mutate(category = case_when(sum_pos != 0 & sum_neg != 0 ~ "Had Some Change",
                               sum_zero ==  sum_pos + sum_neg + sum_zero ~ "Always Ranked the Same",
                            sum_pos == sum_pos + sum_neg + sum_zero ~ paste("Always Ranked Better in", var1),
                            sum_neg == sum_pos + sum_neg + sum_zero ~ paste("Always Ranked Better in", var2 ))) %>%

    ggplot(aes(x = !!sym(glue("{var1}_rank" )), y =!!sym(glue("{var2}_rank" )),
               color  = category,
               text =organization_name
               )) +
    geom_point() +
    geom_function(fun=function(x)x,color="darkred", alpha = .8) +
    labs(x = paste0(var1, " Rank"),
         y =  paste0(var2, " Rank"),
         color = "Consistency of Rankings") +
    theme_c() +
    labs(title = glue("Rank of {var2} vs. Rank of {var1}")) +
      scale_color_brewer(palette ="Set2") +
     facet_wrap(~fiscal_year)+
      theme(plot.title = element_text(size = 14, 
                                      hjust = .5, face="bold",),
            plot.subtitle = element_text(hjust = .5, 
                                         face="italic",
                                         size = 14),
            axis.title = element_text(size = 13, 
                                      face = "bold")) 
  
  ggplotly(plt, margin = m, height = 550) %>% partial_bundle() 

}

plot_ranks_by_consistency <- function(var1, var2, data) {

  
   plt <- data %>%
     filter(fiscal_year > 2010 & fiscal_year < 2021) %>%
    group_by(fiscal_year) %>%
   # arrange(var1) %>%
    mutate("{var1}_rank" := rank(-!!sym(var1)), na.last = "keep") %>%
#    arrange(var2) %>%
    mutate("{var2}_rank"  := rank(-!!sym(var2)),  na.last = "keep")  %>%
     mutate(rank_diff = !!sym(glue("{var2}_rank")) - !!sym(glue("{var1}_rank" ))) %>%
   group_by(EIN) %>%
   mutate(sum_pos = sum(rank_diff >0, na.rm=TRUE),
          sum_neg = sum(rank_diff < 0,  na.rm=TRUE),
          sum_zero = sum(rank_diff ==0, na.rm=TRUE))%>%
     ungroup() %>% 
   mutate(prop_positive = sum_pos / (sum_pos + sum_neg + sum_zero)) %>%
    ggplot(aes(x = !!sym(glue("{var1}_rank" )), y =!!sym(glue("{var2}_rank" )),
               color  = prop_positive,
               text =organization_name
               )) +
    geom_function(fun=function(x)x,color="darkred", alpha = .8, n =201) +
    geom_point() +
    labs(x = paste0(var1, " Rank"),
         y =  paste0(var2, " Rank"),
         title = glue("Rank of {var2} vs. Rank of {var1}"),
         color = glue("Proportion Where {var1}\nRanked Better than\n{var2}")) +
    theme_c(legend.text=element_text(size =8)) +
    scale_color_gradient2(high="#5935CF", low="#D18E01", mid="#E2E2E2", limits=c(0,1), midpoint=.5) +
     facet_wrap(~fiscal_year)+
      theme(plot.title = element_text(size = 12, 
                                      hjust = .5, face="bold",),
            plot.subtitle = element_text(hjust = .5, 
                                         face="italic",
                                         size = 14),
            axis.title = element_text(size = 13, 
                                      face = "bold")) 
  
  ggplotly(plt, margin = m, height = 550) %>% partial_bundle()

}

# 
# print_df_category <- function(data, var1, var2) {
#   data %>%
#      filter(fiscal_year > 2010 & fiscal_year < 2021) %>%
#     group_by(fiscal_year) %>%
#    # arrange(var1) %>%
#     mutate("{var1}_rank" := rank(-!!sym(var1)), na.last = "keep") %>%
# #    arrange(var2) %>%
#     mutate("{var2}_rank"  := rank(-!!sym(var2)),  na.last = "keep")  %>%
#      mutate(rank_diff = !!sym(glue("{var2}_rank")) - !!sym(glue("{var1}_rank" ))) %>%
#    group_by(EIN) %>%
#    mutate(sum_pos = sum(rank_diff >0, na.rm=TRUE),
#           sum_neg = sum(rank_diff < 0,  na.rm=TRUE),
#           sum_zero = sum(rank_diff ==0, na.rm=TRUE))%>%
#      ungroup() %>% 
#    mutate(prop_positive = sum_pos / (sum_pos + sum_neg + sum_zero)) %>%
#     group_by()
# }
# 


# function to plot variables of interest against each other
plot_combo <- function(var1, var2, data) {
  
  data %>%
    ggplot(aes(x = !!sym(var1), y = !!sym(var2), color = EIN)) +
    geom_point(alpha = .9) +
   # geom_line(alpha = .5) +
    facet_wrap(~fiscal_year) +
    viridis::scale_color_viridis(discrete=TRUE,
                                 option = "rocket",
                                 end = .9) +
    theme_bw()+
      theme(plot.title = element_text(size = 18, 
                                      hjust = .5, face="bold",),
            plot.subtitle = element_text(hjust = .5, 
                                         face="italic",
                                         size = 16),
            axis.title = element_text(size = 13, 
                                      face = "bold"),
            legend.position = "none",
            axis.text.x = element_text(angle = 60, vjust = .6)) +
    scale_x_continuous(labels=comma) +
    scale_y_continuous(labels=comma) +
    labs(title = paste0(var2, " vs. ", var1),
         subtitle = "Fill by EIN")
  
}


endowment_data_wide <- endowment_data %>% 
  pivot_wider(names_from=variable_name,
              values_from=value) 



```


# Plotting Endowment Variables Against Each Other, By Year


```{r, eval=TRUE}

vars <-  unique(endowment_data$variable_name)[!grepl("EOY|Admin|Grants", unique(endowment_data$variable_name))]

# pairwise combinations of variables
variable_combinations <- t(combn(vars, 2)) %>%
  as.data.frame()

if (!all_plots) variable_combinations <- variable_combinations[1:4,]


```

```{r endowment by year, results='asis',out.width="60%", eval=FALSE}


cat('## Scale of Original Variables {.tabset}   \n\n')

pwalk(variable_combinations, ~{ 
 cat('### ',paste0(.x, ", ", .y),'\n\n')
 plt <- plot_combo(var1 = .x, var2 = .y, data = endowment_data_wide)
 print(plt)
  cat('\n\n')
 }
)


```

## Where Endowments are Held, By Fiscal Year

```{r, eval=TRUE, results='asis', out.width="100%"}


plt <- endowment_data_wide %>%
  select(contains("EOY"),fiscal_year, 
         EIN, organization_name) %>%
  pivot_longer(cols = contains("EOY"))  %>%
  mutate(name = case_when(
    name == "TermEndowmentBalanceEOYPct" ~ "Temporarily restricted endowment",
    name == "PrmnntEndowmentBalanceEOYPct" ~ "Permanent endowment",
    name == "BoardDesignatedBalanceEOYPct" ~ "Board designated or quasi-endowment"
  )) %>%
  mutate(value = 100*value) %>%
  filter(!is.na(value)) %>%
  ggplot(aes(x=fiscal_year,
             y  = value, 
             color = organization_name)) +
  geom_line() +
  facet_wrap(~name) +
  theme_c(strip.text = element_text(margin =margin(3,0,25,0),
                                    size =4,
                                    lineheight=.5)) +
  scale_color_viridis(option="rocket", discrete= TRUE) +
  labs(y="Percentage of Endowment in Category",
       x = "Fiscal Year") 

ggplotly(plt, margin = m, height =550)  %>% partial_bundle()

```




# Rankings

## Relationships between how companies rank in different variables

```{r, eval=TRUE, results='asis', fig.height = 8, out.width="90%"}



plotlist <- pmap(variable_combinations, ~{ 
 #plt <- plot_ranks(var1 = .x, var2 = .y, data = endowment_data_wide)
   plt <- plot_ranks_by_consistency(var1 = .x, var2 = .y, data = endowment_data_wide) %>% partial_bundle()

 }
)


htmltools::tagList(setNames(plotlist, NULL))


```

## How ranks change over time 

```{r,  results='asis', fig.height = 8, out.width="90%", eval=TRUE}


plot_ranks_over_time <- function(dat,var) {
  plt <- dat %>%
    filter(fiscal_year >=2011 & fiscal_year <=2020) %>%
    group_by(fiscal_year) %>%
    mutate(rank = rank(-!!sym(var), na.last = "keep")) %>%
    filter(!is.na(rank)) %>%
    ggplot(aes(x=fiscal_year, y = rank, color =organization_name)) +
    geom_line() +
    geom_point(alpha = .3, size =.5) +
    scale_color_viridis(option = "rocket", discrete=TRUE) +
    theme_c() +
    scale_y_continuous(n.breaks = 10) +
    scale_x_continuous(breaks=2011:2020) +
    labs(y = paste("Rank of ", var),
         x = "Fiscal Year",
         title = paste0("Rank over time for ", var))
  ggplotly(plt, margin = m, height = 550)
}

plotlist <- map(vars,~ {
  plt<- plot_ranks_over_time(endowment_data_wide,.x) %>% partial_bundle()
})


htmltools::tagList(setNames(plotlist, NULL))





```




# Schedule J 

```{r, eval = FALSE}




m <- list(
    l = 50,
    r = 50,
    b = 50,
    t = 150,
    pad = 0.5
)



source(here("GET_VARS.R"))

files <- dir(here("ballet_990_released_20230208"),
              full.names = TRUE)

get_all_children_by_file(files[5], "/Return/ReturnData/IRS990" )

##################################
# EMPLOYEE INFORMATION
##################################


employee_comp_vars <- c(
  "/Return/ReturnData/IRS990/CYTotalRevenueAmt",
  "/Return/ReturnData/IRS990/TotalEmployeeCnt",
  "/Return/ReturnData/IRS990/EmployeeCnt",
  "/Return/ReturnData/IRS990/CYSalariesCompEmpBnftPaidAmt",
  "/Return/ReturnData/IRS990/CompCurrentOfcrDirectorsGrp/TotalAmt",
  "/Return/ReturnData/IRS990/CompDisqualPersonsGrp/TotalAmt")


employees <- map_df(files, ~get_df(filename = .x, 
                              variables=employee_comp_vars,
                              names=employee_comp_vars)) %>%
  filter_ein()

employees <- employees %>%
  rename_with(~gsub("/Return/ReturnData/IRS990/", "", ., fixed=TRUE)) %>%
  rename(OffDirCompAmt= `CompCurrentOfcrDirectorsGrp/TotalAmt`,
         CompNotOffDir = `CompDisqualPersonsGrp/TotalAmt`)

employees %>%
  select(-filename) %>%
  mutate(across(-c(ReturnTs,EIN, fiscal_year), 
                as.numeric)) %>%
  saveRDS(here("data", "employees.RDS"))




#################
# SCHEDULE J
#################
comp <- map_df(files, ~get_df(filename = .x, schedule = "j"))


comp_clean <- comp %>%
  rename_with(.cols= everything(),
              ~gsub('/Return/ReturnData/IRS990ScheduleJ/', '', .)) %>% 
  select(-contains("Ind")) %>%
  select(fiscal_year, EIN,
         contains("RltdOrgOfficerTrstKeyEmplGrp")) %>%
  # only extract cols within the RltdOrgOfficerTrstKeyEmplGrp
  select(EIN, fiscal_year,
         matches("RltdOrgOfficerTrstKeyEmplGrp\\[.*.\\]/")) %>%
  pivot_longer(-c(EIN,fiscal_year)) %>%
  mutate(id = gsub("\\D", "", name),
       #  name_old = name,
          name = gsub(".*./", "", name),
         id = gsub("990", "", id))

  
comp_clean <- comp_clean %>%
  filter(!is.na(value)) %>%
  distinct() %>% 
  pivot_wider(names_from = name, values_from = value) 



comp_clean <- comp_clean %>%
  mutate(across(contains("Amt"), as.numeric))%>%
  mutate(TitleTxt=tolower(TitleTxt))
  
saveRDS(comp_clean, here("data", "schedj.RDS"))


```

```{r,eval=TRUE}

comp_clean <- read_rds(here("data", "schedj.RDS"))%>%
  left_join(companies_to_ein) %>%
  mutate(fiscal_year = as.numeric(paste(fiscal_year)))

employees <- readRDS(here("data", "employees.RDS")) %>%
    mutate(fiscal_year = as.numeric(paste(fiscal_year)))


```

 
```{r,eval=TRUE}


# clean up title text field because it was free text in the form 990
comp_clean <- comp_clean %>% 
  mutate(TitleTxt = gsub("dancer/choreographer",
                          "dancer / choreographer",
                         TitleTxt),
         TitleTxt = gsub("vp", "Vice President", TitleTxt),
         TitleTxt = gsub("dorector", "director",TitleTxt),
  title_clean = case_when(
    grepl("ceo", TitleTxt, ignore.case = TRUE ) ~"CEO",
    grepl("cfo", TitleTxt, ignore.case = TRUE)~ "Chief Financial Officer",
    grepl("executive dir", TitleTxt, ignore.case = TRUE) ~"Executive Director",
    grepl("artistic dir",TitleTxt,  ignore.case = TRUE) ~"Artistic Director",
    grepl("emeritus|emerita", TitleTxt, ignore.case = TRUE) ~"Emirita/Emiritus Position",
    grepl( "chief dev",TitleTxt, ignore.case=TRUE) &  
    grepl("officer",TitleTxt,  ignore.case = TRUE) ~"Chief Development Officer",
     grepl("director of market|marketing director",TitleTxt,  ignore.case = TRUE) ~ "Director of Marketing",
    grepl("music director",TitleTxt,  ignore.case = TRUE) ~"Music Director",
    grepl("mktg", TitleTxt, ignore.case = TRUE ) & 
      grepl("officer|ofc", TitleTxt, ignore.case = TRUE ) ~ "Marketing Officer",
    grepl("Director of Development",TitleTxt, ignore.case = TRUE) ~ "Director of Development",
    grepl("chief",TitleTxt, ignore.case = TRUE) & 
      grepl("officer",TitleTxt, ignore.case = TRUE) ~ "Other Chief Officer",
    grepl("Dir of Legal",TitleTxt, ignore.case = TRUE) ~"Director of Legal Affairs",
    grepl("Former Senior Dir", TitleTxt, ignore.case = TRUE) ~ "Former Senior Director",
    grepl("Director|Dir", TitleTxt, ignore.case = TRUE) ~ "Other Director",
    grepl("Director", TitleTxt, ignore.case = TRUE) ~ "Other Director",
    TRUE ~ TitleTxt
  )) 

```

## Number of EINs with Each Title


```{r}
# number of EINs with each type of title
comp_clean %>%
  group_by(title_clean) %>%
  summarize(`Number of EINs` = n_distinct(EIN)) %>% 
  arrange(desc(`Number of EINs`))

```


## Number of Individuals with Title 


```{r}
# number of individuals with title
comp_clean %>%
  mutate(title_clean=tolower(title_clean)) %>%
  filter(!is.na(title_clean)) %>%
  group_by(title_clean) %>%
  summarize(`Number of Individuals in Position` = n()) %>% 
  arrange(desc(`Number of Individuals in Position`))

# missingness by variable
# comp_clean %>%
#   select(-c(EIN,fiscal_year,id)) %>%
#   is.na() %>% 
#   colSums() %>%
#   as_tibble(rownames="Variable") %>%
#   mutate(`Not Missing` = nrow(comp_clean) - value) %>%
#   select(-value)


```

## Compensation by Title {.tabset}

### Base Compensation 

```{r, out.width ="80%", eval=TRUE}

comp_clean %>% 
  group_by(title_clean) %>%
  mutate(m = median(BaseCompensationFilingOrgAmt, na.rm= TRUE)) %>%
  filter(!is.na(title_clean)) %>%
  ungroup() %>%
  ggplot(aes(x=fct_reorder(title_clean,m),
             y = BaseCompensationFilingOrgAmt)) +
  geom_jitter(alpha = .5, size = .5, height = 0, width = .05) +
  coord_flip() +
  theme_bw() +
  labs(title = "Base Compensation by Title",
       x = "Title")+
  theme(plot.title = element_text(size = 18, 
                                      hjust = .5, face="bold",),
            plot.subtitle = element_text(hjust = .5, 
                                         face="italic",
                                         size = 16),
            axis.title = element_text(size = 13, 
                                      face = "bold"),
         axis.text.x= element_text(size = 8))


```


### Total Compensation

```{r, out.width ="80%", eval=TRUE}

comp_clean %>% 
  group_by(title_clean) %>%
  mutate(m = median(TotalCompensationFilingOrgAmt, na.rm= TRUE)) %>%
  filter(!is.na(title_clean)) %>%
  ungroup() %>%
  ggplot(aes(x=fct_reorder(title_clean,m),
             y = TotalCompensationFilingOrgAmt)) +
  geom_jitter(alpha = .5, size = .5, height = 0, width = .05) +
  coord_flip() +
  theme_bw() +
  labs(title = "Total Compensation by Title",
       x = "Title")+
  theme(plot.title = element_text(size = 18, 
                                      hjust = .5, face="bold",),
            plot.subtitle = element_text(hjust = .5, 
                                         face="italic",
                                         size = 16),
            axis.title = element_text(size = 13, 
                                      face = "bold"),
         axis.text.x= element_text(size = 8))




```



# Top Employees Compensation Compared to Other Employee Compensation 

* For total employee compensation - `CYSalariesCompEmpBnftPaidAmt: Salaries, other compensation, employee benefits (Part IX, column (A), lines 5–10)`.
* For top employee compensation - Schedule J, looking at total compensation


```{r, eval=TRUE}

top <- comp_clean %>% 
  group_by(EIN,fiscal_year,organization_name) %>% 
  mutate(not_deferred = TotalCompensationFilingOrgAmt -DeferredCompensationFlngOrgAmt) %>%
  summarize(num_top_employees = n(),
            compensation_top_total = sum(TotalCompensationFilingOrgAmt),
            compensation_top_base = sum(BaseCompensationFilingOrgAmt),
            compensation_top_not_def = sum(not_deferred)) %>%
  ungroup()

emp_comp <- employees %>% 
  left_join(companies_to_ein) %>%
  left_join(top) %>%
  mutate(frac = OffDirCompAmt/(CYSalariesCompEmpBnftPaidAmt)) 

```

```{r, eval=TRUE, results='asis',out.width="100%", fig.width = 12}

plt <- emp_comp %>%
  ggplot(aes(x=fiscal_year,
             y = frac,
             color = organization_name,
             group= EIN)) +
  geom_point() +
  geom_line()+
  viridis::scale_color_viridis(discrete=TRUE,
                                 option = "rocket",
                                 end = .9) +
  theme_bw() +
  labs(y="Fraction of Total Compensation Paid",
       title = "Fraction of Total Compensation Paid to C-Suite Employees",
       x="Fiscal Year") +
  scale_y_continuous(n.breaks = 6)+
    theme(plot.title = element_text(size = 14,
                                      hjust = .5, face="bold",),
            plot.subtitle = element_text(hjust = .5,
                                         face="italic",
                                         size = 12),
            axis.title = element_text(size = 13,
                                      face = "bold"))


ggplotly(plt, margin = m, height = 500) %>%
  partial_bundle()



```


```{r,eval=TRUE}

plt <- emp_comp %>%
  ggplot(aes(x=fiscal_year,
             y = compensation_top_total/CYSalariesCompEmpBnftPaidAmt,
             color = organization_name,
             group= EIN)) +
  geom_point() +
  geom_line()+
  viridis::scale_color_viridis(discrete=TRUE,
                                 option = "rocket",
                                 end = .9) +
  theme_bw() +
  labs(y="Fraction of Total Compensation Paid",
       title = "Fraction of Total Compensation Paid to Employees Reported in Schedule J",
       x="Fiscal Year") +
  scale_y_continuous(n.breaks = 6)+
    theme(plot.title = element_text(size = 14,
                                      hjust = .5, face="bold",),
            plot.subtitle = element_text(hjust = .5,
                                         face="italic",
                                         size = 12),
            axis.title = element_text(size = 13,
                                      face = "bold"))


ggplotly(plt, margin = m, height = 500) %>%
  partial_bundle()


```


```{r, eval=FALSE}

 plt <- emp_comp %>%
   mutate(num_not_top = TotalEmployeeCnt - num_top_employees,
          compensation_not_top = CYSalariesCompEmpBnftPaidAmt - compensation_top_total,
          avg_not_top = compensation_not_top/num_not_top ,
          avg_top = compensation_top_total / num_top_employees) %>%
   filter(fiscal_year !=2014 & fiscal_year !=2021) %>% 
   # something strange with Aspen Santa Fe 
   filter(avg_not_top > 0) %>% 
  ggplot(aes(y=avg_not_top, 
             x =avg_top,
             label=fiscal_year,
             color =organization_name,
             group = EIN)) +
  geom_point() +
   theme_c() + 
  viridis::scale_color_viridis(discrete=TRUE,
                                 option = "rocket",
                                 end = .9) +
    theme(plot.title = element_text(size = 12, 
                                      hjust = .5, face="bold",
                                      margin = margin(5,5,5,5)),
            plot.subtitle = element_text(hjust = .5, 
                                         face="italic",
                                         size = 14),
            axis.title = element_text(size = 13, 
                                      face = "bold"),
          axis.text.x = element_text(angle = 20, vjust = .6)) +
   scale_y_continuous(labels =comma) +
   labs(y = "Average Employee Compensation",
        x = "Average C-Suite Compensation",
        title = "Average C-Suite Pay versus Average Employee Compensation") +
   facet_wrap(~fiscal_year) +
   scale_x_continuous(labels = comma)
 
ggplotly(plt, margin = marg, height = 500) %>%
   partial_bundle()




```


```{r}
marg <- list(
    l = 50,
    r = 50,
    b = 250,
    t = 250,
    pad = 0.5
)
 
 

plt <-  emp_comp %>%
   mutate(num_not_top = TotalEmployeeCnt - num_top_employees,
          compensation_not_top = CYSalariesCompEmpBnftPaidAmt - compensation_top_total,
          avg_not_top = compensation_not_top/num_not_top ,
          avg_top = compensation_top_total / num_top_employees) %>%
   group_by(fiscal_year) %>%
   mutate(rank_avg_not_top = rank(-avg_not_top,  na.last ="keep"),
          rank_avg_top = rank(-avg_top, na.last="keep")) %>%
   filter(fiscal_year !=2014 & fiscal_year !=2021) %>%
   ggplot(aes(x =rank_avg_top, y = rank_avg_not_top, color = organization_name)) +
   geom_point() +
   theme_c()+
   labs(y = "Ranking of Average Employee Compensation",
        x = "Ranking of Average C-Suite Compensation",
        title = "Average C-Suite Pay versus Average Employee Compensation") +
   facet_wrap(~fiscal_year)  +
  geom_function(fun=function(x)x,color="darkred", alpha = .8) +
    viridis::scale_color_viridis(discrete=TRUE,
                                 option = "rocket",
                                 end = .9) 
 
 ggplotly(plt, margin = marg, height = 500) %>%
   partial_bundle()

 ```
 
 
 
 
 
 ```{r,eval=TRUE}
 
 ranks <- emp_comp %>%
   mutate(num_not_top = TotalEmployeeCnt - num_top_employees,
          compensation_not_top = CYSalariesCompEmpBnftPaidAmt - compensation_top_total,
          avg_not_top = compensation_not_top/num_not_top ,
          avg_top = compensation_top_total / num_top_employees,
          organization_name = ifelse(is.na(organization_name), EIN, organization_name)) %>%
   group_by(fiscal_year) %>%
   mutate(rank_avg_not_top = rank(-avg_not_top,  na.last ="keep"),
          rank_avg_top = rank(-avg_top, na.last="keep")) %>%
   filter(fiscal_year !=2014 & fiscal_year !=2021) %>%
   mutate(rank_diff = rank_avg_top - rank_avg_not_top ) %>%
   group_by(organization_name) %>%
   mutate(sum_pos = sum(rank_diff >0, na.rm=TRUE),
          sum_neg = sum(rank_diff < 0,  na.rm=TRUE),
          sum_zero = sum(rank_diff ==0, na.rm=TRUE)) %>%
   select(fiscal_year, rank_avg_not_top, rank_avg_top, contains("sum"), organization_name) %>%
   ungroup() %>%
   mutate(category = case_when(sum_pos != 0 & sum_neg != 0 ~ "Had Some Change",
                               sum_zero ==  sum_pos + sum_neg + sum_zero ~ "Always Ranked the Same",
                            sum_pos == sum_pos + sum_neg + sum_zero ~  "Always Ranked Better in Average Employee Compensation" ,
                            sum_neg == sum_pos + sum_neg + sum_zero ~ "Always Ranked Better in C-Suite Compensation"),
          prop_positive = sum_pos/(sum_pos+sum_zero+sum_neg)) 
 
 
plt <- ranks %>%
   ggplot(aes(x =rank_avg_top, y = rank_avg_not_top, 
              label = organization_name,
              color = prop_positive)) +
   geom_point() +
   theme_c()+
   labs(y = "Ranking of Average Employee Compensation",
        x = "Ranking of Average C-Suite Compensation",
        title = "Rankings of Average C-Suite Pay versus Average Employee Compensation",
        color = "Proportion Where C-Suite Compensation\nWas Ranked Better Than Average Employee Compensation") +
   facet_wrap(~fiscal_year)  +
  geom_function(fun=function(x)x,color="darkred", alpha = .8) +
 # scale_fill_distiller(palette ="PuOr", low="#EFC950", mid="#E2E2E2") +
  scale_color_gradient2(high="#5935CF", low="#D18E01", mid="#E2E2E2", limits=c(0,1), midpoint=.5)
 
 ggplotly(plt, margin = marg, height = 500) %>%
   partial_bundle()
 
 

```


```{r}

 emp_comp %>%
   mutate(num_not_top = TotalEmployeeCnt - num_top_employees,
          compensation_not_top = CYSalariesCompEmpBnftPaidAmt - compensation_top_total,
          avg_not_top = compensation_not_top/num_not_top ,
          avg_top = compensation_top_total / num_top_employees) %>%
  filter(!is.na(avg_not_top) & !is.na(avg_top)) %>%
  group_by(EIN) %>%
  mutate(min_year = min(fiscal_year)) %>%
  filter(fiscal_year == min_year | fiscal_year == 2020) %>%
    arrange(fiscal_year) %>%
  mutate(pct_change_comp_top = ifelse(
      lag(avg_top, n =1) != 0,
      (avg_top - lag(avg_top, n =1))/
             lag(avg_top, n =1),
      NA),
          pct_change_comp_not_top  = ifelse(
      lag(avg_not_top, n =1) != 0,
      (avg_not_top - lag(avg_not_top, n =1))/
             lag(avg_not_top, n =1),
      NA)) %>%
  filter(fiscal_year == 2020) %>%
   mutate(diff =pct_change_comp_top - pct_change_comp_not_top) %>%
  left_join(companies_to_ein) %>%
  filter(!is.na(diff)) %>%
  ggplot(aes(x =fct_reorder(organization_name, diff), y = diff)) +
  geom_bar(stat="identity") +
  coord_flip() +
  theme_c() +
  labs(title = "Percent Change in Average Executive Compensation Minus Percent Change in Average Employee Compensation",
       y = "",
       x = "Difference in Percent Changes")



 emp_comp %>%
   mutate(num_not_top = TotalEmployeeCnt - num_top_employees,
          compensation_not_top = CYSalariesCompEmpBnftPaidAmt - compensation_top_total,
          avg_not_top = compensation_not_top/num_not_top ,
          avg_top = compensation_top_total / num_top_employees) %>%
  filter(!is.na(avg_not_top) & !is.na(avg_top)) %>%
  group_by(EIN) %>%
  mutate(pct_change_comp_top = ifelse(
      lag(avg_top, n =1) != 0,
      (avg_top - lag(avg_top, n =1))/
             lag(avg_top, n =1),
      NA),
          pct_change_comp_not_top  = ifelse(
      lag(avg_not_top, n =1) != 0,
      (avg_not_top - lag(avg_not_top, n =1))/
             lag(avg_not_top, n =1),
      NA)) %>%
   ungroup() %>%
   filter(organization_name != "Aspen Santa Fe Ballet") %>%
   select(pct_change_comp_not_top, pct_change_comp_top, 
          organization_name, EIN, organization_name, fiscal_year) %>%
   ggplot(aes(x= pct_change_comp_not_top, y = pct_change_comp_top))+
   geom_point() +
   facet_wrap(~name)



```



