---
title: "Generate Table from XML Files"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{css,echo=FALSE}

/* CSS for formatting of the knitted html file; changing default colors of readthedown */

#sidebar h2 { background: #3B5561 !important;}
#main .nav-pills > li.active > a,
#main .nav-pills > li.active > a:hover,
#main .nav-pills > li.active > a:focus {
    background-color: #22983B;
}

#main .nav-pills > li > a:hover {
    background-color: #22983B;
}

// font color to dark blue 
h1, h2, h3, h4, h5, h6, legend {
    color: #334B56;
}


#table-of-contents h2 {
    background-color: #3B5561;
}

/* expand content to screenwidth rather than cutting it off */
#content{
max-width:2300px;
}



```


```{r}

library(tidyverse)
library(xml2)
library(kableExtra)
library(lubridate)
library(forcats)
library(here)


```

```{r}
all_data_filtered <- read_rds(here("data", "data_filter_990.RDS"))
```

```{r}

# make kable table with consistent formatting
make_table <- function(..., title = "") {
  title <- paste0("<center><span style = 'font-size:150%;color:black'><b>",
                  title,
                  "</span></b><center>")
   as_tibble(...) %>%
    kbl(caption = title) %>%
    kable_material() %>%
    row_spec(row=0, background = "#43494C" , color = "white", bold = TRUE)
}

```



# 1. Filings per Year

## Filings by Fiscal Year

```{r}

############################################################
# TABLE: Number of Filings by Year of Return Date
############################################################
all_data_filtered %>%
  group_by(fiscal_year) %>%
  summarize(`Number of Filings`= n()) %>%
  rename(`Year of Return Date` = fiscal_year) %>%
  make_table(title = "Number of Filings by Fiscal Year")


```

```{r, fig.width = 10, fig.height = 6}

############################################################
# PLOT: Number of Filings by Year of Return Date
############################################################
all_data_filtered %>%
  group_by(fiscal_year) %>%
  summarize(`Number of Filings`= n()) %>%
  rename(`Fiscal Year` = fiscal_year) %>%
  ggplot(aes(x =`Fiscal Year`, y = `Number of Filings`)) +
  geom_bar(stat = "identity", fill = "#3D6071") +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(size = 15),
        axis.text.y=element_text(size = 12)) +
  labs(title = "Number of Filings by Fiscal Year") +
  scale_y_continuous(n.breaks = 6)


```

# 2. Companies per Year

```{r, fig.width = 10, fig.height = 6}


####################################################
# TABLE OF (DISTINCT) EINS PER YEAR
####################################################
all_data_filtered %>%
  group_by(fiscal_year) %>%
  summarize(`Number of Companies` = n_distinct(EIN)) %>%
  rename(`Fiscal Year` = fiscal_year) %>%
  make_table(title = "Number of Companies by Fiscal Year")


####################################################
# PLOT OF (DISTINCT) EINS PER YEAR
####################################################
all_data_filtered %>%
  group_by(fiscal_year) %>%
  summarize(`Number of Distinct EINs` = n_distinct(EIN)) %>%
  rename(`Fiscal Year` = fiscal_year) %>%
  ggplot(aes(x = `Fiscal Year`, y = `Number of Distinct EINs`)) +
  geom_bar(stat = "identity", fill = "#3D6071") +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        axis.text.x = element_text(size = 15)) +
  labs(title = "Number of Companies by Fiscal Year") 
     


```

# 3. Companies with More than One Filing in a Year

We can see the EINs and names of the companies with more than one filing in some fiscal year in the table below.

```{r}
############################################################
# TABLE WITH COMPANIES WITH MORE THAN ONE FILING, BY YEAR
############################################################
all_data_filtered %>%
  group_by(fiscal_year, EIN) %>%
  # number of filings for each company for each year
  summarize(n_filings =n(),.groups="drop") %>%
  filter(n_filings > 1) %>%
  rename(`Fiscal Year` = fiscal_year,
         `Number of Filings` = n_filings) %>%
  arrange( `Fiscal Year`, desc(`Number of Filings`)) %>%
  make_table(
    title = "Companies with More than One Filing<br>by Fiscal Year") %>%
  scroll_box(height = "450px")

```

<br><br><br>


We can also look at the trends in the number of companies with 1 filing or 2 filings by  fiscal year in the plot below.


```{r, fig.width = 12, fig.height = 6}


all_data_filtered %>%
  group_by(fiscal_year, EIN) %>%
  # number of filings for each company for each year
  summarize(n_filings =n(),.groups="drop") %>% 
  group_by(n_filings, fiscal_year) %>%
  # number of companies with n_filings for each year
  summarize(n_companies = n(), .groups="drop") %>%
  mutate(n_filings = factor(n_filings)) %>%
  ggplot(aes(x = n_filings, 
             y =n_companies,
             fill = n_filings)) +
  geom_bar(stat = "identity",
           position = "dodge",
           show.legend=FALSE) +
  facet_wrap(~fiscal_year, nrow =1) +
  scale_fill_manual(
    values = c("#72e5ef",
               "#1e438d", 
               "#58df8c")) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 15),
        strip.text = element_text(face="bold",size = 18)) +
  labs(y = "Number of Companies",
       x = "Number of Filings",
       title = "Companies by Number of Filings per Year",
       subtitle = "By Fiscal Year")

```


# 4. Gaps in Filings 

## Gaps in Filings for All Companies {.tabset}

### By EIN 

```{r, fig.height = 16,  fig.width = 10}

all_data_filtered %>% 
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time))%>%
  ggplot(aes(xmin = TaxPeriodBeginDt, 
             xmax = TaxPeriodEndDt, 
             y = fct_reorder(EIN, sum_coverage))) +
  geom_linerange(size = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 4),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "EIN",
       x = "Date",
       title = "Gaps in Coverage by EIN")
  # scale_x_date(date_breaks = "9 months",
  #              date_labels = "%b %Y") +

```

### By Company Name


```{r, fig.height = 16,  fig.width = 10}


ein_to_name <- all_data_filtered %>%
  select(EIN, BusinessName) %>%
  group_by(EIN) %>%
  mutate(n_names = n_distinct(BusinessName)) %>%
  slice_head(n=1) %>%
  select(EIN, BusinessName)


all_data_filtered %>% 
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time))%>%
  left_join(ein_to_name) %>%
  ggplot(aes(xmin = TaxPeriodBeginDt, 
             xmax = TaxPeriodEndDt, 
             y = fct_reorder(BusinessName, sum_coverage))) +
  geom_linerange(size = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 4),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "Business Name",
       x = "Date",
       title = "Gaps in Coverage by Business Name") 
  # scale_x_date(date_breaks = "9 months",
  #              date_labels = "%b %Y") +

```


## Companies with at Least One Gap Year

To check for gap years, we can take the difference between `TaxPeriodBeginDt` and the lagged `TaxPeriodEndDt` for each EIN.

```{r}

###################################################
# TABLE WITH COMPANIES WITH AT LEAST ONE GAP YEAR
###################################################
all_data_filtered %>% 
  # do not include ammended returns for this computation
  filter(is.na(AmendedReturnInd) & ReturnTypeCd != "990T") %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  group_by(EIN) %>%
  arrange(TaxPeriodBeginDt) %>%
  mutate(before =  lag(TaxPeriodEndDt, n =1),
        gap = interval(before, TaxPeriodBeginDt),
        gap = as.numeric(gap, "days")) %>%
  filter(gap >=365)  %>% 
  select(EIN,
         `Previous End of Fiscal Year` = before,
         `Beginning of Fiscal Year` = TaxPeriodBeginDt,
         `Gap Between Filings (Days)` = gap) %>%
  make_table(title = "Companies with a Gap of One Year or More Between Filings") %>%
  scroll_box(height = "450px")

# vector of EINs corresponding to companies with at least one gap year
ein_with_gap <- all_data_filtered %>% 
  # do not include ammended returns for this computation
  filter(is.na(AmendedReturnInd) & ReturnTypeCd != "990T") %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  group_by(EIN) %>%
  arrange(TaxPeriodBeginDt) %>%
  mutate(before =  lag(TaxPeriodEndDt, n =1),
        gap = interval(before, TaxPeriodBeginDt),
        gap = as.numeric(gap, "days")) %>%
  filter(gap >=365) %>%
  pull(EIN)

```


### Plotting Gaps in Coverage for Companies with at Least One Gap Year {.tabset}
 
#### By EIN

```{r, fig.width = 11}

############################################################
# PLOT COVERAGE FOR COMPANIES WITH 1+ GAP YEAR (BY EIN)
############################################################
all_data_filtered %>% 
  filter(EIN %in% ein_with_gap) %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time))%>%
  ggplot(aes(xmin = TaxPeriodBeginDt, 
             xmax = TaxPeriodEndDt, 
             y = fct_reorder(EIN, sum_coverage))) +
  geom_linerange(size = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 8),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "EIN",
       x = "Date",
       title = "Gaps in Coverage for Companies\nwith at Least One Gap Year")

# gap_cvg_basicplot <- all_data_filtered
#   ggplot(aes(xmin = TaxPeriodBeginDt, 
#              xmax = TaxPeriodEndDt, 
#              y = fct_reorder(EIN, sum_coverage))) +
#   geom_linerange(size = 1, alpha = .8) +
#   theme_bw() +
#   theme(plot.title = element_text(size = 16, face = "bold", hjust = .5),
#         axis.title = element_text(size = 16, face = "bold"),
#         plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
#         axis.text.x = element_text(size = 11,angle=20, vjust = .7),
#         axis.text.y = element_text(size = 8),
#         strip.text = element_text(face="bold",size = 18)) +
#   scale_x_date(date_breaks = "12 months",
#                date_labels = "%b %Y") +
#   labs(y = "EIN",
#        x = "Date",
#        title = "Gaps in Coverage for Companies\nwith at Least One Gap Year")
# 
# gap_cvg_basicplot
# ggsave("gap_cvg_basicplot.png", path = here("final_report", "images"))

```

#### By Business Name

```{r, fig.width = 11}

############################################################
# PLOT COVERAGE FOR COMPANIES WITH 1+ GAP YEAR (BY NAME)
############################################################
all_data_filtered %>% 
  filter(EIN %in% ein_with_gap) %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time))%>%
  left_join(ein_to_name) %>%
  ggplot(aes(xmin = TaxPeriodBeginDt, 
             xmax = TaxPeriodEndDt, 
             y = fct_reorder(BusinessName, sum_coverage))) +
  geom_linerange(size = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 16, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 8),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "EIN",
       x = "Date",
       title = "Gaps in Coverage for Companies\nwith at Least One Gap Year")

```


## Distribution of Gaps Between Filings

```{r}
#####################################
# HISTOGRAM OF GAPS BETWEEN FILINGS
#####################################
all_data_filtered %>% 
  # do not include ammended returns for this computation
  filter(is.na(AmendedReturnInd) & ReturnTypeCd != "990T") %>%
  # compute coverage so we can order the plot by days of total coverage by EIN
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  group_by(EIN) %>%
  arrange(TaxPeriodBeginDt) %>%
  mutate(before =  lag(TaxPeriodEndDt, n =1),
        gap = interval(before, TaxPeriodBeginDt),
        gap = as.numeric(gap, "days")) %>%
  ggplot(aes(x =gap )) +
  geom_histogram( fill = "#3D6071") +
  theme_bw() +
  labs(x = "Gap Between Filings (Days)",
       y = "Number of Companies",
       title = "Distribution of the Gap Between Filings") +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11, vjust = .7),
        axis.text.y = element_text(size = 12),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_continuous(n.breaks = 10)





```

