---
title: "Endowment Management"
date: "Last updated on `r Sys.Date()`"
author: "Ruiz-Fuentes"
format: html
editor: visual
theme: cosmo
number-sections: true
---

This section will demonstrate calculations and exploratory data visualizations on the endowment management and governance using Form 990, Section D, Part V.

By investigating variables 1a-g we can learn about the **financial health** and **performance** of the endowment fund including some of its sources of revenue, expense and impacts of the investment activities. This will also produce insights on the endowment fund's **long-term sustainability** and **perpetuity** which is especially meaningful for smaller endowment's because they can learn from the successes of a larger endowment. It may also help them reevaluate their endowment management and how governance alters in the face economic downturns and market volatility.

```{r, include=FALSE}
## Set default behavior for all code chunks here:
knitr::opts_chunk$set(
  echo = TRUE, 
  message = FALSE, 
  warning = FALSE, 
  fig.width = 16/2, 
  fig.height = 9/2)
```

```{r, data wrangling, include=FALSE}

## Load packages
library(dplyr) # data manipulation
library(readr) # scrapes web pages
library(ggplot2) # data visualization
library(janitor) # data manipulation

## Set directort
setwd("/Users/michelruiz-fuentes/Desktop/smith-capstone-24/990s_assetallocation")

## Load data
schedule_d <- read.csv("/Users/michelruiz-fuentes/Desktop/smith-capstone-24/990s_assetallocation/schedule_d_vals.csv")  
companies <- read.csv("/Users/michelruiz-fuentes/Desktop/smith-capstone-24/990s_assetallocation/companies.csv") 

# Cleaning column names
schedule_d <- schedule_d %>%
  clean_names()

schedule_d_cleaned <- 
  schedule_d %>% 
  rename(
    administrative_expenses = x_return_return_data_irs990schedule_d_cy_endwmt_fund_grp_administrative_expenses_amt, 
    boy_balance = 
x_return_return_data_irs990schedule_d_cy_endwmt_fund_grp_beginning_year_balance_amt,
    contributions = 
x_return_return_data_irs990schedule_d_cy_endwmt_fund_grp_contributions_amt,
    eoy_balance=
x_return_return_data_irs990schedule_d_cy_endwmt_fund_grp_end_year_balance_amt,
    grants_or_scholarships =
x_return_return_data_irs990schedule_d_cy_endwmt_fund_grp_grants_or_scholarships_amt,
    investment_earnings_or_losses = x_return_return_data_irs990schedule_d_cy_endwmt_fund_grp_investment_earnings_or_losses_amt,
    other_expenditure = 
x_return_return_data_irs990schedule_d_cy_endwmt_fund_grp_other_expenditures_amt)

# Changing EIN to character - schedule_d
schedule_d_cleaned <- schedule_d_cleaned %>%
  mutate(fiscal_year = as.character(fiscal_year)) 

# Changing EIN to character - schedule_d
schedule_d_cleaned <- schedule_d_cleaned %>%
  mutate(ein = as.character(ein)) 

# Changing EIN to character - companies
companies <- companies %>%
  mutate(ein = as.character(ein)) %>%
  select(ein, organization_name)

# new df with focusedcolumns
schedule_d_cleaned <- schedule_d_cleaned %>% 
  select("ein", 
      "fiscal_year",
      "administrative_expenses", 
      "boy_balance", 
      "contributions", 
      "eoy_balance", 
      "grants_or_scholarships", 
      "investment_earnings_or_losses", 
      "other_expenditure")

# Joining datasets 
schedule_d_cleaned <- schedule_d_cleaned %>%
  left_join(companies , by = 'ein', relationship = "many-to-many")

# Converting data from wide to long and selecting Part V columns from 990s
schedule_d_2017 <- subset(schedule_d_cleaned, fiscal_year == "2017")
schedule_d_2018 <- subset(schedule_d_cleaned, fiscal_year == "2018")
schedule_d_2019 <- subset(schedule_d_cleaned, fiscal_year == "2019")
schedule_d_2020 <- subset(schedule_d_cleaned, fiscal_year == "2020")
schedule_d_2021 <- subset(schedule_d_cleaned, fiscal_year == "2021")
schedule_d_2022 <- subset(schedule_d_cleaned, fiscal_year == "2022")
```

### About the Data \[[1](https://www.irs.gov/pub/irs-pdf/i990sd.pdf)\]

`Schedule_d_cleaned` contains 1061 observations and 10 variables. Three of the variables are characters: the organization name, fiscal year and EIN code; while seven of the variables are integers demonstrating different quantitative aspects of the endowment fund.

For example,

-   `organization_name` : Company names provided by DDP

-   `ein` : unique nine-digit number assigned by the IRS to identity a business entity, 501(c) non-profits

-   `fiscal_year` : a 12-month period that a non-profit organization uses for financial reporting and budgeting purposes

-   `boy_balance` **\[1a\]**: beginning-of-year balances on the organization's endowment funds for the current year; includes cumulative total of funds and assets held in the endowment from previous periods

-   `contributions` **\[1b\]**: enter the amounts of contributions; includes donor gifts, grants, monetary donations and contributions in the form of assets or securities, as well as additional funds established by the organization's governing board to function like an endowment

-   `investment_earnings_or_losses` **\[1c\]**: represents income or losses generates from investing the endowments funds in financial markets (such as interest, dividends, capital gains or losses on investments), including both realized and unrealized amounts; for example reported net of transaction costs and earnings reported on a gross basis

-   `grants_or_scholarships` **\[1d\]**: enter the amounts distributed for grants and scholarships; scholarships represent direct aid to individuals, they are distinguished from general programmatic aid referenced; funds disbursed to support specific programs, projects or individuals

-   `other_expenditure` **\[1e\]**: encompasses any additional expenses incurred by the endowment fund during the fiscal year that are not classified as administrative expenses or grants/scholarships; may include fees for investment management, legal expenses, or other operational costs

-   `administrative_expenses` **\[1f\]**: expenses can arise from either internal or third-party sources; relate to general operation and management such as salaries of administrative staff, office rent, utilities, and other overhead expenses

-   `eoy_balance` **\[1g\]**: determined by adding investing earnings and subtracting investment losses to the beginning of year balance, after grants or scholarships have been awarded and other expenditures

### Change Over Fiscal Year

How does it evaluate the fund's [financial health]{.underline}: a healthy endowment may have consecutive positive changes in beginning of year balances from end of year balances; indicating a stronger endowment value attributed to larger contributions or investment earnings

How does it evaluate the fund's [sustainability]{.underline} : trends of these endowment balances and their differences over multiple fiscal years can inform long-term planning and decision-making; perhaps making adjustments to total expenses or strategizing to increase investment earnings or contributions for the endowment's continued growth and perpetuity

```{r}
# creating a column with the difference between eoy - boy
# calculating the change in endowment balance
schedule_d_cleaned <- schedule_d_cleaned %>%
  mutate(difference = (eoy_balance - boy_balance))
```

```{r, include=FALSE}
# new df with diffs and vars of interest
change_in_amt_df <- schedule_d_cleaned %>%
  select(organization_name, fiscal_year, boy_balance, eoy_balance, difference)
```

```{r, include=FALSE}
# Converting data from wide to long and selecting Part V columns from 990s
change_in_amt_df_2017 <- subset(change_in_amt_df, fiscal_year == "2017")
change_in_amt_df_2018 <- subset(change_in_amt_df, fiscal_year == "2018")
change_in_amt_df_2019 <- subset(change_in_amt_df, fiscal_year == "2019")
change_in_amt_df_2020 <- subset(change_in_amt_df, fiscal_year == "2020")
change_in_amt_df_2021 <- subset(change_in_amt_df, fiscal_year == "2021")
change_in_amt_df_2022 <- subset(change_in_amt_df, fiscal_year == "2022")
```

```{r}
# example snapshot of change in fiscal year balance
sanfrancisco_diff <- change_in_amt_df %>%
  filter(organization_name == "San Francisco Ballet")
# print the first few rows 
print(head(sanfrancisco_diff))
```

### Expense Ratio

How does it evaluate the fund's [financial health]{.underline}: a healthy endowment may have a decreasing expense ratio because the firm may become more efficient at allocating expenses while an increase may indicate a financial strain or higher expenses on the firm.

How does it evaluate the fund's [sustainability:]{.underline} a consistent expense ratio may reveal a pattern into the endowment's operations and their spending habits - maintaining a 'sustainable' ratio is crucial to maintaining the endowment's perpetuity and preserving their abilities to meet their objectives

```{r, include=FALSE}
# creating a column titled total_expenses
# sum of administrative_expenses, grants_or_scholarships, other_expenditure
schedule_d_cleaned <- schedule_d_cleaned %>%
  mutate(total_expenses = rowSums(
    select(., administrative_expenses, grants_or_scholarships, other_expenditure)))
```

```{r}
# calculating expense ratio
# assess proportion of total expenses relative to the endowment's total assets
schedule_d_cleaned <- schedule_d_cleaned %>%
  mutate(expense_ratio = (total_expenses / eoy_balance) * 100)
```

```{r, include=FALSE}
# new df with expense ratios and vars of interests
expense_ratio_df <- schedule_d_cleaned %>%
  select(organization_name, fiscal_year, expense_ratio, eoy_balance, total_expenses)
```

```{r, include=FALSE}
# Converting data from wide to long and selecting Part V columns from 990s
expense_ratio_df_2017 <- subset(expense_ratio_df, fiscal_year == "2017")
expense_ratio_df_2018 <- subset(expense_ratio_df, fiscal_year == "2018")
expense_ratio_df_2019 <- subset(expense_ratio_df, fiscal_year == "2019")
expense_ratio_df_2020 <- subset(expense_ratio_df, fiscal_year == "2020")
expense_ratio_df_2021 <- subset(expense_ratio_df, fiscal_year == "2021")
expense_ratio_df_2022 <- subset(expense_ratio_df, fiscal_year == "2022")
```

```{r}
# example snapshot of expense ratios for SF
sanfrancisco_er <- expense_ratio_df %>%
  filter(organization_name == "San Francisco Ballet")
# print the first few rows 
print(head(sanfrancisco_er ))
```

```{r, include=FALSE}
expense_ratio_spread17 <- 
  ggplot (expense_ratio_df_2017, aes(x= organization_name, y= expense_ratio)) +
  geom_boxplot () + 
  facet_wrap(~eoy_balance, scales = "free") + 
  labs(title = "Distribution of Expense Ratios",
    subtitle = "Schedule D, Part V, 2017 Filings",
    x = "Expense Ratio (%)",
    y = "Frequency") 
```

```{r}
expense_ratio_spread17
```

### Change in Endowment Balance

### 
