---
title: "Retrieve Endowment Funds"
author: "Rose Evard"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css', 'template.css')
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "output_html")})
---

```{r, message = FALSE}

library(tidyverse)
library(kableExtra)
library(lubridate)
library(forcats)
library(here)
library(viridis)

knitr::opts_chunk$set(message = FALSE, warning = FALSE)

```

```{r}

#######################################
# CREATE BIG PALETTE WITH MANY COLORS
#######################################

pal1 <- viridis_pal(option="mako", end = .9)(12)
pal2 <- viridis_pal(option="rocket", end = .8)(12)
pal3 <- viridis_pal(option="magma", end = .9)(12)
pal4 <- viridis_pal(option="inferno", end = .8)(12)

set.seed(123)
pal <- c(pal1, pal2, pal3, pal4)
indexes <- sample.int(length(pal), replace=FALSE)
pal <- pal[indexes]
```

```{r}

# make kable table with consistent formatting
make_table <- function(..., title = "", col_names = c("")) {
  title <- paste0("<center><span style = 'font-size:150%;color:black'><b>",
                  title,
                  "</span></b><center>")
   as_tibble(...) %>%
    kbl(caption = title,
        col.names = col_names) %>%
    kable_material() %>%
    row_spec(row=0, background = "#43494C" , color = "white", bold = TRUE)
}

```

> Note: Zero was coded as false and One as true for Checkbox questions.  
> Additional note: Put the companies.csv file into your data folder 

```{r retrieving endowment info}
endowment_data <- read_rds(here("data", "endowment_filter_data_990.RDS"))
names <- readRDS(here("data","companies.RDS"))
names <- names %>% 
  rename(organization_name = Company)
names <- names %>%
  mutate(organization_name = ifelse(
    organization_name == "Ballet Hispanico", 
    "Ballet Hispánico",
    organization_name))

```

```{r}
##Graphing difference between beginning and end periods  

endowment_data %>%
  select(ReturnDate, EIN, CYBeginningYearBalanceAmt, CYEndYearBalanceAmt) %>%
  pivot_longer(cols = c(CYBeginningYearBalanceAmt, CYEndYearBalanceAmt), names_to = "Period", values_to = "Balance") %>%
  mutate(Period = as.factor(Period)) %>%
  mutate(Year = as.integer(substr(ReturnDate, 1, 4))) %>%
  filter(Year <= 2022) %>% # Filtering to include only data up to 2022
  ggplot(aes(x = Period, y = Balance, group = EIN)) +
  facet_wrap(~Year) +
  geom_point(alpha = 0.5) + 
  geom_line(alpha = 0.5) + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 35, hjust =1),
        legend.position = "none") +
  labs(title = "Comparing Fiscal Year Beginning and End Endowment Totals") +
  scale_y_continuous(labels = scales::label_dollar())

```





## Which Years and Companies Filled out Schedule D  

```{r}
##Identifying those that DID fill out schedule D for Endowment, as they supplied info in any numeric column
##Assuming the sum of from cols 4-31 (schedule D finances) is not zero
filed_schd_d <- endowment_data %>%
 mutate(Filled_Sch_D = ifelse(rowSums(across(CYBeginningYearBalanceAmt:TermEndowmentBalanceEOYPct), na.rm = TRUE) != 0, "Yes", "No")) %>%
  select(TaxPeriodBeginDt, TaxPeriodEndDt, ReturnDate, EIN, DonorRstrOrQuasiEndowmentsInd, Filled_Sch_D) %>%
  mutate(Need_Sch_D = case_when(is.na(DonorRstrOrQuasiEndowmentsInd) ~ "NA",
                                DonorRstrOrQuasiEndowmentsInd == TRUE ~ "Yes",
                                DonorRstrOrQuasiEndowmentsInd == FALSE ~ "No")) %>%
  select(-DonorRstrOrQuasiEndowmentsInd)
```

```{r}
filed_schd_d %>%
  select(EIN, Filled_Sch_D) %>%
  filter(Filled_Sch_D == "Yes") %>%
  group_by(EIN) %>%
  summarize(num_times_fil_sche_D = n()) %>%
  left_join(names, by = "EIN") %>%
  arrange(desc(num_times_fil_sche_D)) %>%
  select(EIN, organization_name, num_times_fil_sche_D) %>%
  make_table(title = "EINs that reported an endowment at least once", col_names = c("EIN", "Organization Name", "# Times an endowment was filled out")) %>%
  scroll_box(height = "450px")

```

```{r}
# plot to show same information as previous table
filed_schd_d %>%
  select(EIN, Filled_Sch_D) %>%
  filter(Filled_Sch_D == "Yes") %>%
  group_by(EIN) %>%
  summarize(num_times_fil_sche_D = n()) %>%
  group_by(num_times_fil_sche_D) %>%
  summarize(num_ein = n()) %>%
  ggplot(aes(x = num_times_fil_sche_D, y = as.integer(num_ein))) +
  geom_bar(stat="identity") +
  scale_x_continuous(breaks = 1:7) +
  labs(title = "For Companies that Reported and Endowment Least Once,\nNumber of Endowment Reports",
       x = "Number of Endowment Reports",
       y = "Number of EINs") +
  geom_label(aes(label = num_ein)) +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = .5, face="bold"),
        plot.subtitle = element_text(hjust = .5, face="italic"),
        axis.text.x = element_text(size = 13),
        axis.title = element_text(size = 16, face = "bold"),
        legend.position = "none")
```

There are `r filed_schd_d %>% filter(Filled_Sch_D == "Yes") %>%pull(EIN) %>% unique() %>% length()` EINs which, at one point, have filled out Schedule D. 

```{r}
#How many unique schedule D's are there
filed_schd_d %>%
 select(EIN, Filled_Sch_D) %>%
  filter(Filled_Sch_D == "Yes") %>%
  summarize(num_sch_d = n()) %>%
  make_table(title = "Total Number of Unique Endowment Reports", col_names = c("Total"))  %>%
  row_spec(row = 0:1, align = "center")
```

There are 288 uniquely filed Schedule D's. 

```{r}
filed_schd_d %>%
  select(TaxPeriodEndDt, EIN, Filled_Sch_D) %>%
  filter(Filled_Sch_D == "Yes") %>%
  mutate(year = substr(TaxPeriodEndDt, 1, 4)) %>%
  group_by(year) %>%
  summarize(num_sch_d_in_year = n()) %>%
  arrange(desc(year)) %>%
  make_table(title = "Years with Endowment Reported", col_names = c("Year", "Total Endowments Reported filed within the year"))
```

Years represented in the data are 2016-2023, based on the beginning financial year date. Since there are only 9 reported endowment in 2023, our analysis will focus on 2016-2022.


## Determining Which Needed to File Schedule D for Endowment

Not all those that fill out schedule D do so for endowment, so I focused on Question 10 "Did the organization, directly or through a related organization, hold assets in temporarily restricted endowments, permanent endowments, or quasi endowments?  If "yes," complete Schedule D, Part V".  

```{r}
##Counting those that needed to fill schedule D
filed_schd_d %>%
  group_by(Need_Sch_D) %>%
  summarize(n = n()) %>%
  arrange(desc(Need_Sch_D)) %>%
  make_table(title = "Total EINs checked YES on Q10", col_names = c("Need Schedule D on Q10", "Total Filings"))
```

```{r}
# plot showing same information as previous table
filed_schd_d %>%
  group_by(Need_Sch_D) %>%
  summarize(`Total Filings` = n()) %>%
  arrange(desc(Need_Sch_D)) %>%
  ggplot(aes(x = Need_Sch_D, y = `Total Filings`)) +
  geom_bar(stat="identity") +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = .5, face="bold"),
        plot.subtitle = element_text(hjust = .5, face="italic"),
        axis.text.x = element_text(size = 15, face="bold"),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 16, face = "bold")) +
  labs(x = "Response on Q10",
       title = "Responses for Whether They Needed Schedule D on Q10") +
  scale_y_continuous(n.breaks = 6)
```

```{r}

##Seeing those that DID and or DID NOT need to fill out Schedule D, yet Did or Did Not
filed_schd_d %>%
  group_by(Need_Sch_D, Filled_Sch_D) %>%
  summarize(n = n()) %>%
  arrange(desc(Need_Sch_D)) %>%
  make_table(title = "Whether Individual 990 Filings Reported An Endowment", col_names = c("Needed to Report an Endowment", "Reported an Endowment", "Total Reports"))



```


```{r, fig.width = 10}
# bar plot with same information as table above
filed_schd_d %>%
  group_by(Need_Sch_D, Filled_Sch_D) %>%
  summarize(n = n()) %>%
  arrange(desc(Need_Sch_D)) %>%
  ungroup() %>%
  mutate(Need_Sch_D = paste0("Needed to Report Endowment: ", Need_Sch_D)) %>%
  ggplot(aes(x = Filled_Sch_D, y = n)) +
  geom_bar(stat="identity") +
  facet_wrap(~Need_Sch_D) +
  theme_bw() +
  theme(plot.title = element_text(size = 16, hjust = .5, face="bold"),
        plot.subtitle = element_text(hjust = .5, face="italic"),
        axis.text.x = element_text(size = 15, face="bold"),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 16, face = "bold"),
        strip.text = element_text(face="bold",size = 14)) +
  labs(title = "Number of Filings where Endowments were Reported",
       subtitle = "By Whether They Actually Needed it by Question 10",
       x = "Endowment Reported",
       y = "Number of Filings")
```



Total which needed schedule D and proceeded to file it matches total who filed it previously.  

# Examining Gaps in Filing Schedule D's  
         
```{r}

filed_schd_d %>%
  filter(Filled_Sch_D == "Yes") %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  ggplot(aes(xmin = as.Date(TaxPeriodBeginDt, "%Y-%m-%d"), 
             xmax = TaxPeriodEndDt, 
             y = EIN)) +
  geom_linerange(linewidth = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 4),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "EIN",
       x = "Date",
       title = "Endowments reported by EIN\nacross time")
```

```{r}
#Getting EINS of those who have filled out Schedule D
ein_with_sch_d <- filed_schd_d %>%
  filter(Filled_Sch_D == "Yes") %>%
  pull(EIN) %>%
  unique()

#Gap plot
filed_schd_d %>%
  filter(EIN %in% ein_with_sch_d) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  ggplot(aes(xmin = as.Date(TaxPeriodBeginDt, "%Y-%m-%d"), 
             xmax = TaxPeriodEndDt, 
             y = EIN)) +
  geom_linerange(linewidth = 1, alpha = .8) +
  theme_bw() +
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 4),
        strip.text = element_text(face="bold",size = 18)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  labs(y = "EIN",
       x = "Date",
       title = "Filing history for all EINs\nthat reported an endowment at least once")
```


```{r}
#Retrieving list of schedule D filing gaps 
filed_schd_d %>%
  filter(EIN %in% ein_with_sch_d & Filled_Sch_D == "Yes") %>%
  select(EIN, 
         TaxPeriodBeginDt, #calc dates b/w schedule d 
         TaxPeriodEndDt) %>%
  group_by(EIN) %>%
  arrange(TaxPeriodBeginDt) %>%
  mutate(before =  lag(TaxPeriodEndDt, n =1),
        gap = interval(before, TaxPeriodBeginDt),
        gap = as.numeric(gap, "days")) %>%
  filter(gap > 1) %>%
  left_join(names, by = "EIN") %>%
  select(EIN, organization_name, gap, before, TaxPeriodBeginDt) %>%
  make_table(title = "Endowment Reporting Gaps", col_names = c("EIN", "Organization Name", "Gap Length", "Gap Start Date", "Gap End Date"))
```

```{r}
#Retrieving list of 990 filing gaps, for those who have once filled out an EIN
filed_schd_d %>%
  filter(EIN %in% ein_with_sch_d) %>%
  select(EIN, 
         TaxPeriodBeginDt, 
         TaxPeriodEndDt) %>%
  group_by(EIN) %>%
  arrange(TaxPeriodBeginDt) %>%
  mutate(before =  lag(TaxPeriodEndDt, n =1),
        gap = interval(before, TaxPeriodBeginDt),
        gap = as.numeric(gap, "days")) %>%
  filter(gap > 1) %>%
  left_join(names, by = "EIN") %>%
  select(EIN, organization_name, gap, before, TaxPeriodBeginDt) %>%
  make_table(title = "990 Filing Gaps for EINs\nwhich once reported an Endowment", col_names = c("EIN", "Organization Name", "Gap Length", "Gap Start Date", "Gap End Date"))
```

# Aligning 990 Filing Gaps and Schedule D Filing Gaps  
Important note: Filing Schedule D guarantees they filed a 990.  


```{r}
##Plotting both 990 gaps and Schedule D Gaps
schd_d_gap <- filed_schd_d %>%
  filter(Filled_Sch_D == "Yes") %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time),
         type = "Endowment Reported")

##Gap plot  
gap_plot_fr <- filed_schd_d %>%
  filter(EIN %in% ein_with_sch_d) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time),
         type = "Only 990 Filed") %>%
  bind_rows(schd_d_gap) %>%
  left_join(names, by = "EIN") %>%
  ggplot(aes(xmin = as.Date(TaxPeriodBeginDt, "%Y-%m-%d"), 
             xmax = TaxPeriodEndDt, 
             y = organization_name,
             color = type)) +
  geom_linerange(linewidth = 1, alpha = .8) +
  theme_bw() + 
  theme(plot.title = element_text(size = 20, face = "bold", hjust = .5),
        axis.title = element_text(size = 16, face = "bold"),
        plot.subtitle = element_text(size = 16, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 11,angle=20, vjust = .7),
        axis.text.y = element_text(size = 4),
        strip.text = element_text(face="bold",size = 18),
        legend.text = element_text(size = 5)) +
  scale_x_date(date_breaks = "12 months",
               date_labels = "%b %Y") +
  scale_color_manual(values = c("blue4", "darkgoldenrod1")) +
  labs(y = "Company",
       x = "Fiscal Year",
       title = "Endowment Reporting & 990 Filing",
       color = "Type of Filing")
gap_plot_fr

## Table Identifying those that don't line up 
filed_schd_d %>%
  filter(EIN %in% ein_with_sch_d) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time),
         type = "Only 990 Filed") %>%
  bind_rows(schd_d_gap) %>%
  group_by(EIN) %>%
  arrange(EIN, type, TaxPeriodBeginDt) %>%
  select(-type) %>%
  anyDuplicated()  

filed_schd_d %>%
  filter(EIN %in% ein_with_sch_d) %>%
  mutate(time = interval(TaxPeriodBeginDt, TaxPeriodEndDt),
         time = as.numeric(time, "days")) %>%
  group_by(EIN) %>%
  mutate(sum_coverage = sum(time),
         type = "Only 990 Filed") %>%
  bind_rows(schd_d_gap) %>%
  select(-c(type, time, sum_coverage)) %>%
  group_by(TaxPeriodBeginDt, TaxPeriodEndDt, ReturnDate, EIN, Filled_Sch_D, Need_Sch_D) %>%
  mutate(num_dups = n()) %>%
  filter(num_dups < 2) %>%
  left_join(names, by = "EIN") %>%
  group_by(EIN, organization_name) %>%
  summarize() %>%
  make_table(title = "EINS that filled out Schedule D sometimes,\nbut not always", col_names = c("EIN", "Organization Name"))

```
There are seven EINs which do not always fill out Schedule D when they fill out 990.  

# Checking Endowment Reporting - Is the Math Right?  

```{r}
## Using Quinn's endowment data 
endowment_data <- read_rds(here("data", "endowments_by_most_recent_filings.RDS"))
```

```{r}
## NOTE:  ASSUMING NA IS ZERO FOR THE SAKE OF THE CALCULATIONS
# Finding differences in calculation and report

endowment_data$BeginningYearBalanceAmt <- as.numeric(as.character(endowment_data$BeginningYearBalanceAmt))
endowment_data$ContributionsAmt <- as.numeric(as.character(endowment_data$ContributionsAmt))
endowment_data$OtherExpendituresAmt <- as.numeric(as.character(endowment_data$OtherExpendituresAmt))
endowment_data$AdministrativeExpensesAmt <- as.numeric(as.character(endowment_data$AdministrativeExpensesAmt))
endowment_data$GrantsOrScholarshipsAmt <- as.numeric(as.character(endowment_data$GrantsOrScholarshipsAmt))
endowment_data$InvestmentEarningsOrLossesAmt <- as.numeric(as.character(endowment_data$InvestmentEarningsOrLossesAmt))
endowment_data$EndYearBalanceAmt <- as.numeric(as.character(endowment_data$EndYearBalanceAmt))

test_diff <- endowment_data %>%
  mutate(across(.cols = 3:9, ~ ifelse(is.na(.), 0, .))) %>%
  mutate(end_bal_calc = (BeginningYearBalanceAmt + 
           ContributionsAmt + InvestmentEarningsOrLossesAmt - 
           abs(OtherExpendituresAmt) - abs(AdministrativeExpensesAmt) - abs(GrantsOrScholarshipsAmt)),
         same_end = ifelse(end_bal_calc == EndYearBalanceAmt, TRUE, FALSE),
         diff_end = EndYearBalanceAmt - end_bal_calc)

## MEAGHAN FIX THIS CODE BLOCK
## as_tibble validate vs name_repair error

# Looking at differences
# test_diff %>%
#   filter(same_end == FALSE) %>%
# #  arrange(desc(abs(diff_end))) %>%
#   left_join(names, by = "EIN") %>%
#   select(organization_name, fiscal_year, EndYearBalanceAmt, end_bal_calc, diff_end) %>%
#   make_table("Differences in Reported End Year Balances", 
#              col_names = c("Organization Name", "Year", "Reported End Balance", 
#                            "Calculated End Balance", "Difference")) %>%
#   scroll_box(height = "450px")

## Misreports by company
test_diff %>%
  filter(same_end == FALSE) %>%
  left_join(names, by = "EIN") %>%
  ggplot(aes(x = organization_name)) + 
  geom_bar(aes(fill = organization_name), show.legend = FALSE) + 
  theme_bw() + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 90),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  scale_fill_viridis_d() +
  labs( title = "Number of Times Companies Miscalculated End Year Balance",
        y = "Number of Miscalculations",
        x = "Company Name")

## Proportion of misreports by company 
##List of companies
companies_miscalculate <- test_diff %>%
  filter(same_end == FALSE) %>%
  left_join(names, by = "EIN") %>%
  group_by(organization_name) %>%
  summarize(n = n()) %>%
  arrange(n) %>%
  pull(organization_name) %>%
  unique()

## Counts per company
counts <- test_diff %>%
  left_join(names, by = "EIN") %>%
  filter(same_end == FALSE) %>%
  group_by(organization_name) %>%
  summarize(n = n())

hist_diff_fr <- test_diff %>%
  left_join(names, by = "EIN") %>%
  mutate(organization_name = factor(organization_name, level = companies_miscalculate)) %>%
  filter(organization_name %in% c(companies_miscalculate)) %>%
  ggplot(aes(x = organization_name, fill = same_end)) + 
  geom_bar() +
  theme_bw() + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7),
        legend.title = element_text(size = 7)) + 
  scale_fill_viridis_d(labels = c("Yes", "No")) +
  labs( title = "Number of Times Companies Miscalculated End Year Balance",
        y = "Total Reports",
        x = "Company Name",
        fill = "Was End Balance \nMiscalculated?") +
  coord_flip()
hist_diff_fr

#Histograms of all misreports compared (sorta)
test_diff %>%
  filter(same_end == FALSE) %>%
  left_join(names, by = "EIN") %>%
  select(EIN, fiscal_year, organization_name, EndYearBalanceAmt, end_bal_calc) %>%
  pivot_longer(cols = c(EndYearBalanceAmt, end_bal_calc),
               names_to = "type",
               values_to = "balance") %>%
  mutate(type = ifelse(type == "EndYearBalanceAmt", "Reported", "Calculated")) %>%
  ggplot(aes(x = type, y = balance, fill = type)) +
  geom_col(position = "dodge") +
  facet_wrap(~organization_name, scales = "free") + 
  theme_bw() + 
    theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) +
  scale_y_continuous(labels = scales::label_dollar()) +
  scale_fill_manual(values = c("darkgoldenrod1", "blue4"))


# Difference Plotted
calc_diff <- test_diff %>%
  filter(same_end == FALSE) %>%
  left_join(names, by = "EIN") %>%
  select(EIN, fiscal_year, organization_name, EndYearBalanceAmt, diff_end) %>%
  mutate(name = paste(organization_name, fiscal_year, sep = ", ")) %>%
  ggplot(aes(y = name, x = diff_end, fill = organization_name)) +
  geom_col(position = "dodge", show.legend = FALSE) +
  #facet_wrap(~organization_name, scales = "free") + 
  theme_bw() + 
    theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 5),
        axis.text.y = element_text(size = 5),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) +
  scale_x_continuous(labels = scales::label_dollar()) + 
 # scale_fill_manual(values = pal)+ 
  labs(title = "Difference Between Calculated and Reported End Balances",
       x = "Calculated Difference",
       y = "Company Name and Fiscal Year") + 
  geom_text(aes(label = scales::dollar(diff_end)), size = 2)
calc_diff


# Relationship between year and misreports?  No
test_diff %>%
  filter(same_end == FALSE) %>%
  left_join(names, by = "EIN") %>%
  ggplot(aes(x = fiscal_year)) + 
  geom_bar() + 
  theme_bw() + 
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10, angle = 90),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) + 
  labs( title = "Miscalculations per Year End Year Balance")

##MEAGHAN FIX THIS CODE BLOCK
## same problem as above

## Table showing same information as plot above, total miscalculations
# test_diff %>%
#   filter(same_end == FALSE) %>%
#   arrange(desc(abs(diff_end))) %>%
#   left_join(names, by = "EIN") %>%
#   group_by(organization_name) %>%
#   summarize(n = n()) %>%
#   make_table("List of Misreported Companies", 
#              col_names = c("Organization Name", "Total Miscalculations")) %>%
#   scroll_box(height = "450px")
```

# Saving Relevant Images For Report 

## Schedule D Gap Plot  

```{r}
gap_plot_fr
ggsave("gap_plot_endowment.png", path = here("final_report", "images"))
```


## Reporting Consistencies

```{r}
hist_diff_fr
ggsave("miscalc_end_bal.png", path = here("final_report", "images"))
```

```{r}
calc_diff
ggsave("diff_end_bal.png", path = here("final_report", "images"))
```




