---
title: "Unpaid Labor Analysis - Evard"
author: "Rose Evard"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css', 'template.css')
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "output_html")})
---

```{r setup, include=FALSE}
library(here)
library(tidyverse)
library(kableExtra)
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

# make kable table with consistent formatting
make_table <- function(..., title = "", col_names = c("")) {
  title <- paste0("<center><span style = 'font-size:150%;color:black'><b>",
                  title,
                  "</span></b><center>")
   as_tibble(...) %>%
    kbl(caption = title,
        col.names = col_names) %>%
    kable_material() %>%
    row_spec(row=0, background = "#43494C" , color = "white", bold = TRUE)
}

```

```{r}
source(here("GET_VARS.R"))
```

Volunteer variables:  

      Return//ReturnData//<VotingMembersGoverningBodyCnt>13</VotingMembersGoverningBodyCnt>
      <VotingMembersIndependentCnt>13</VotingMembersIndependentCnt>
      <TotalEmployeeCnt>77</TotalEmployeeCnt>
      <TotalVolunteersCnt>13</TotalVolunteersCnt>
      
```{r}
files <- dir( here("ballet_990_released_20230208"),
              full.names = TRUE)  
names <- read_csv(here("data", "companies.csv")) %>% 
  mutate(EIN = as.character(ein)) %>%
  select(-ein)
##Specifically reading in volunteer data
volun_data <- map_df(files, ~
                    get_df(variables =  c("//Return//ReturnData//TotalEmployeeCnt", 
                                          "//Return//ReturnData//TotalVolunteersCnt"),
                           filename = .x
                           ))

## Wrangling, adding organization name 
volun_data<- volun_data %>%
  left_join(names, by = "EIN") %>%
  mutate(TotalEmployeeCnt = as.numeric(TotalEmployeeCnt),
         TotalVolunteersCnt = as.numeric(TotalVolunteersCnt))
```

## Basic Stats about Volunteers 
```{r}
ggplot(volun_data, aes(x = fiscal_year, y = TotalVolunteersCnt)) + 
  geom_point() + 
  theme_bw()

##histogram
ggplot(volun_data, aes(x = TotalVolunteersCnt)) + 
  geom_histogram(binwidth = 20) + 
  theme_bw()

## How many didn't answer question to  volunteers 
volun_data %>%
  mutate(report_volun = ifelse(!is.na(TotalVolunteersCnt), "Answer", "NA")) %>%
  group_by(report_volun) %>%
  summarize(n = n()) %>%
  make_table(title = "How many 990s ANSWER volunteer question", col_names = c("Answer Question", "Total 990s"))

## How many report no volunteers
volun_data %>%
  filter(TotalVolunteersCnt < 1) %>%
  group_by(TotalVolunteersCnt) %>%
  summarize(n = n())
```

## Basic Stats about Employees
```{r}
ggplot(volun_data, aes(x = fiscal_year, y = TotalEmployeeCnt)) + 
  geom_point() + 
  theme_bw()

##histogram
ggplot(volun_data, aes(x = TotalEmployeeCnt)) + 
  geom_histogram(binwidth = 20) + 
  theme_bw()

## How many 990s that don't report anything 
volun_data %>%
  mutate(report_employ = ifelse(!is.na(TotalEmployeeCnt), "Yes", "No")) %>%
  group_by(report_employ) %>%
  summarize(num_990 = n())

##Count of NAs 
count_NA_employs <- volun_data %>% 
  mutate(report_employ = ifelse(!is.na(TotalEmployeeCnt), "Yes", "No")) %>% 
  filter(report_employ == "No") %>% 
  group_by(EIN) %>% 
  summarize(n = n()) %>% 
  nrow() 

## How many NA reports per company
volun_data %>%
  mutate(report_employ = ifelse(!is.na(TotalEmployeeCnt), "Yes", "No")) %>%
  filter(report_employ == "No") %>%
  group_by(EIN) %>%
  summarize(report_per_company = n()) %>%
  arrange(desc(report_per_company))

## How many reports didn't have any employees??? 
volun_data %>%
  filter(TotalEmployeeCnt < 1) %>%
  group_by(TotalEmployeeCnt) %>%
  summarize(num_990 = n())

volun_data %>%
  filter(TotalEmployeeCnt < 1) %>%
  group_by(EIN) %>%
  summarize(company_total = n()) %>%
  arrange(desc(company_total))

count_no_employs <- volun_data %>%
  filter(TotalEmployeeCnt < 1) %>%
  group_by(EIN) %>%
  summarize(company_total = n()) %>%
  nrow()
```

There are `r count_NA_employs` companies who have missing data regarding employees.  

There are `r count_no_employs` companies who reported zero employees.  

> Do those who don't report employees report volunteers?  
> What's the revenue for those companies?  
> Start doing company-by-company analysis 
