---
title: "Building Endowment Preliminary Analysis"
author: "Zhen Nie"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "./output_html")})
---
```{r library, message =FALSE, warning=FALSE, echo = FALSE}
library(dplyr)
library(readr)
library(wesanderson)
library(purrr)
library(xml2)
library(tidyverse)
library(ggplot2)
library(kableExtra)
library(plotly)
library(shiny)
library(here)

```

```{r}
#functions of making graphs and tables
make_graph <- function(df){
  graph <- ggplot(df, aes(x = year, y = BookValueSum, color = as.factor(BusinessName))) +
  geom_line() +
  theme_bw() +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10),
        strip.text = element_text(face="bold",size = 5))
  return(graph)
}


```


## Importing variables related to the endowment of building 

```{r, warning=FALSE }
# use get_var
files<- dir(here("990archivesfeb2024"),
              full.names = TRUE)  
source(here::here("GET_VARS.R"))

suppressMessages(buildings <- map_df(files, ~get_df(filename = .x , 
                              schedule = 'd')) %>% 
  filter_ein())

for (i in colnames(buildings)) {
  if (i != "ReturnTs" & i != "EIN" & i!= "filename") {
    New_name = gsub('/Return/ReturnData/IRS990ScheduleD/', '', i)
    buildings = plyr::rename(buildings, setNames(New_name, i))
  }} 
select(buildings, ReturnTs, EIN,OtherLandBuildingsGrp,TotalBookValueLandBuildingsAmt,BuildingsGrp,fiscal_year) 



```


```{r}
companies <- read_csv(here("data/companies.csv")) %>%
  mutate(EIN = as.character(EIN))

#load in regional information
# add state information to buildings dataframe
ein_to_state <- readRDS(here("./data/ein_to_state.RDS")) 

test_state<-buildings %>% 
  left_join(y = ein_to_state, by = "EIN") %>% 
  mutate(
    TotalBookValueLandBuildingsAmt = as.numeric(TotalBookValueLandBuildingsAmt)
  )

test_state <- unique(test_state) %>% 
  mutate(name = organization_name)
test_state <- unique(test_state)
```

## Preliminary Analysis of PartIV Data

Term explained: 
*Book value* is an accounting term used for both a measure of a business's equity and the value of an asset as it appears on a balance sheet. As time goes on, the cost stays the same, but the accumulated depreciation increases, so the book value decreases.
```{r}
# filter out all data each year
data_each_year<-list(
  data_2017 <- filter(buildings, grepl("2017", ReturnTs)),
  data_2018 <- filter(buildings, grepl("2018", ReturnTs)),
  data_2019 <- filter(buildings, grepl("2019", ReturnTs)),
  data_2020 <- filter(buildings, grepl("2020", ReturnTs)),
  data_2021 <- filter(buildings, grepl("2021", ReturnTs)),
  data_2022 <- filter(buildings, grepl("2022", ReturnTs)),
  data_2023 <- filter(buildings, grepl("2023", ReturnTs))
)

```



```{r}
BookValueSum <- list()

# restore each year's sum for each EIN's buildings bookvalue 
for(i in c(1:7)){
  data = data_each_year[[i]]
  df <- data.frame(data %>% 
    filter(!is.na(TotalBookValueLandBuildingsAmt)) %>% 
    mutate_at(c("TotalBookValueLandBuildingsAmt"), as.numeric) %>% 
    group_by(EIN) %>% 
    summarise(
      BookValueSum = sum(TotalBookValueLandBuildingsAmt)
    ) %>% 
    mutate(year = 2016+i,
           EIN = as.numeric(EIN)))
           
  #print(df)
  BookValueSum[[i]] = df
}
#  convert the list into dataframe
bookvalues <- do.call(rbind, BookValueSum)
```


```{r mutate bookvalue dataframe with quantile and business name}
# rank book value based on quantile

bookvalues<-bookvalues %>%
   mutate(quantile = ntile(BookValueSum, 10),
          EIN = as.character(EIN)) %>% 
   arrange(desc(EIN)) %>% 
   left_join(y = companies, by = "EIN") %>% 
   left_join(y = ein_to_state, by = "EIN") %>% 
   distinct()

bookvalues[267, "organization_name"] = "Mystic Ballet INC"
bookvalues[268, "organization_name"] = "Mystic Ballet INC"
bookvalues[269, "organization_name"] = "Mystic Ballet INC"
bookvalues[270, "organization_name"] = "Mystic Ballet INC"
bookvalues[271, "organization_name"] = "Mystic Ballet INC"
bookvalues[272, "organization_name"] = "Mystic Ballet INC"
bookvalues[273, "organization_name"] = "Mystic Ballet INC"

#rank of BVS 
ranking_BVS<-bookvalues %>%
  arrange(desc(BookValueSum)) %>%  
  select(EIN, organization_name, year, BookValueSum) %>% 
  kbl(caption = "Ranking of Building Book Values ") %>%
  kable_material() %>%
  row_spec(row = 0, background = "red", color = "white", bold = TRUE)
ranking_BVS

bookvalues<- unique(bookvalues)
saveRDS(bookvalues,here('data','bookvalues.RDS'))
```


```{r divide bookvalues into high medium and low}
#dataframe with higher Book Values
high <- bookvalues %>% 
  filter(year == 2020) %>% 
  filter(quantile >= 8) %>% 
  select(EIN) %>% 
  inner_join(bookvalues, by = "EIN") %>% 
   select(EIN, BookValueSum, year, quantile, organization_name) #organization_name.x

medium <- bookvalues %>% 
  filter(year == 2020) %>% 
  filter(quantile > 3 & quantile < 7) %>% 
  select(EIN) %>% 
  inner_join(bookvalues, by = "EIN") %>% 
   select(EIN, BookValueSum, year, quantile, organization_name)
  
low <- bookvalues %>% 
  filter(year == 2020) %>% 
  filter(quantile <= 3) %>% 
  select(EIN) %>% 
  inner_join(bookvalues, by = "EIN") %>% 
   select(EIN, BookValueSum, year, quantile, organization_name)

```
 
```{r}
# visualization of largest 10 book values
#Largest_10 <- bookvalues
ranking_10<-bookvalues %>%
  arrange(desc(BookValueSum)) %>%  
  select(EIN, organization_name, year, BookValueSum) %>% 
  head(10) 

ranking_10_vis <- ggplot(ranking_10, aes(x = as.factor(organization_name), y = BookValueSum, color = as.factor(year), fill = as.factor(year))) +
  geom_bar(binwidth = 0.15, position="dodge", stat='identity', alpha=0.3) +
  geom_text(aes(label = year), fontface = "bold", vjust = 1.5,
             position = position_dodge(0.9), size = 4) +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) +
  labs(y = "Schedule D Sum of Buildings Book Value ($)",
       x = "Year",
       title = "Largest 10 Schedule D Sum of Buildings Book Values",
       subtitle = "By Fiscal Year")
ggplotly(ranking_10_vis)
```

```{r visualizatoins of the change in buildings book value}
# visualization of overall book value
total <- ggplot(bookvalues, aes(x = year, y = BookValueSum, color = as.factor(organization_name), fill = as.factor(organization_name))) +
  geom_bar(binwidth = 0.15, position="dodge", stat='identity', alpha=0.3) +
  #theme_minimal() + 
  #scale_color_brewer(palette="BrBG")+
  #theme_bw()+
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 12, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10),
        strip.text = element_text(face="bold",size = 5),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7)) +
  labs(y = "Schedule D Sum of Buildings Book Value ($)",
       x = "Year",
       title = "Schedule D Sum of Buildings Book Value Change",
       subtitle = "By Fiscal Year")
ggplotly(total)
```


```{r}
# separate EIN with medium and low building book value according to previous graph
# high
high <- ggplot(high, aes(x = year, y = BookValueSum, color = as.factor(organization_name))) +
  geom_line()+
  theme_bw() +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 8, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 7),
        strip.text = element_text(face="bold",size = 5)) +
  labs(y = "Schedule D Sum of Buildings Book Value",
       x = "Year",
       title = "Schedule D Sum of Buildings Book Value Change (High)",
       subtitle = "By Fiscal Year")

ggplotly(high)


```


```{r}
# medium
medium <- ggplot(medium, aes(x = year, y = BookValueSum, color = as.factor(organization_name))) +
  geom_line()+
  theme_bw() +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 8, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 7),
        strip.text = element_text(face="bold",size = 5)) +
  labs(y = "Schedule D Sum of Buildings Book Value",
       x = "Year",
       title = "Schedule D Sum of Buildings Book Value Change (Medium)",
       subtitle = "By Fiscal Year")

ggplotly(medium)
```


```{r}
# low
  ggplot(low, aes(x = year, y = BookValueSum, color = as.factor(organization_name))) +
  geom_line()+
  theme_bw() +
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10),
        strip.text = element_text(face="bold",size = 5)) +
  labs(y = "Schedule D Sum of Buildings Book Value",
       x = "Year",
       title = "Schedule D Sum of Buildings Book Value Change (Low)",
       subtitle = "By Fiscal Year")

```
The book value will decrease because of depreciation, so the normal trend of an EIN from previous visualization should be a slight downward line. I will filter out EIN with a bizarre trend that should be :
  1) Abruptly going down
  2) going up at any point
  3) appear less than 3 years

## Unusual EIN's
```{r}
# rank book value based on quantile
EIN_appearl_less_than_3<-bookvalues %>%
  mutate(quantile = ntile(BookValueSum, 10),
         EIN = as.character(EIN)) %>% 
  group_by(EIN) %>% 
  summarise(appear_times = n()) %>% 
  filter(appear_times < 3) %>% 
  inner_join(y=companies, by = c("EIN")) %>% 
  kbl(caption = "Unusual EIN (appear less than 3 times)") %>%
  kable_material() %>%
  row_spec(row = 0, background = "red", color = "white", bold = TRUE)
EIN_appearl_less_than_3
```


## Percentage change of building's book value
```{r}
changes <- bookvalues %>% 
  group_by(EIN) %>% 
  mutate(change = (BookValueSum - lag(BookValueSum)) / lag(BookValueSum) * 100) %>%
  select(EIN, organization_name, year, BookValueSum, change, quantile) %>% 
  mutate(change_abs = abs(change)) 

high_changes <- changes %>% 
  filter(year == 2020) %>% 
  filter(quantile >= 8) %>% 
  select(EIN) %>% 
  inner_join(changes, by = "EIN") %>% 
   select(EIN, BookValueSum, year, quantile, change, organization_name) #organization_name.x

medium_changes <- changes %>% 
  filter(year == 2020) %>% 
  filter(quantile > 3 & quantile < 7) %>% 
  select(EIN) %>% 
  inner_join(changes, by = "EIN") %>% 
   select(EIN, BookValueSum, year, quantile, change, organization_name)
  
low_changes <- changes %>% 
  filter(year == 2020) %>% 
  filter(quantile <= 3) %>% 
  select(EIN) %>% 
  inner_join(changes, by = "EIN") %>% 
   select(EIN, BookValueSum, year, quantile, change, organization_name)

#Making an infinite value NA
changes[145, ]$change = NA

```

```{r}
# line graph of all percentage change from 2017-2022
all_changes_pct <- ggplot(changes, aes(x = year, y = changes$change, color =as.factor(organization_name), fill =  as.factor(organization_name))) +
  geom_bar(position="dodge", stat='identity', alpha=0.3) +
  #theme_bw() +
  theme(legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7))+
  theme(plot.title = element_text(size = 13, face = "bold", hjust = .5),
        axis.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 10, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        strip.text = element_text(face="bold",size = 10)) +
  labs(y = "Change in Buildings Book Value Each Year (%)",
       x = "Year",
       title = "% Change in Bookvalues of Building Endowment each Fiscal Year",
       color = "EIN") 

ggplotly(all_changes_pct)
```


```{r}
# all_changes_pct <- ggplot(changes, aes(x = year, y = changes$change, color =  as.factor(organization_name), fill = as.factor(organization_name))) +
#   geom_bar(position="dodge", stat='identity', alpha=0.3) +
#   theme_bw() +
#   theme(legend.key.size = unit(1, 'mm'),
#         legend.text = element_text(size=7))+
#   theme(plot.title = element_text(size = 6, face = "bold", hjust = .5),
#         axis.title = element_text(size = 4, face = "bold"),
#         plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
#         axis.text.x = element_text(size = 4),
#         axis.text.y = element_text(size = 4),
#          strip.text = element_text(face="bold",size = 5)) +
#    labs(y = "Change in Buildings Book Value Each Year (%)",
#         x = "Year",
#         title = "% Change in Bookvalues of Building Endowment each Fiscal Year",
#         color = "EIN") + 
#     facet_wrap(~quantile_exp)
#  
# ggplotly(all_changes_pct)


```


```{r}
# filter out EIN with normal depreciation change
filtered_changes <- changes %>% 
  filter(change_abs > 10) %>% 
  ggplot(aes(x = year, y = change, color = as.factor(organization_name))) +
  geom_line(na.rm = FALSE, size = 0.2) +
  ylim(-200, 800)+
  theme_bw() +
  theme(legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7))+
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 6),
        strip.text = element_text(face="bold",size = 5)) +
  labs(y = "Change in Buildings Book Value Each Year (%)",
       x = "Year",
       title = "Change in Bookvalues of Building Endowment each Fiscal Year",
       color = "EIN")


ggplotly(p = filtered_changes)
```


```{r}
big_change<-changes %>% 
  filter(change_abs > 10) %>% 
  kbl(caption = "Changes larger than 10%") %>%
  kable_material() %>%
  row_spec(row = 0, background = "red", color = "white", bold = TRUE)
big_change  
```


## Book Value Changes %

```{r}
# summarize all companies changes
summary(changes$change) 
```

```{r}
#ranking of all change in book values
ranking<-changes %>%
  arrange(desc(change_abs)) %>% 
  select(EIN, organization_name, year, change, BookValueSum) %>% 
  kbl(caption = "Ranking of Change in Book Value Per Year  (%)") %>%
  kable_material() %>%
  row_spec(row = 0, background = "red", color = "white", bold = TRUE)
ranking
```


```{r}
#filter out small changes 
change_less_than_10<-changes %>% 
  filter(change_abs <=10)

#istogram of all large changes 
ggplotly(
  ggplot(change_less_than_10, aes(x = change_abs)) + 
  geom_histogram()+
  theme_bw()
  )
```


```{r}
# filter out all % change larger than 10
change_larger_than_10 <- changes %>% 
  filter(change_abs > 10)

# generate histogram of all absolute changes 
ggplotly(
  ggplot(change_larger_than_10, aes(x = change_abs)) + 
  geom_histogram()+
  theme_bw()
  )

```


```{r}
# All summary statistics 
changes %>% 
  drop_na() %>% 
  group_by(year) %>% 
  summarise(mean = mean(change),
            median = median(change),
            max = max(change),
            min = min(change)) %>% 
  kbl(caption = "Summary statistics of Building Book Value Changes (Fiscal Year)") %>%
  kable_material() %>%
  row_spec(row = 0, background = "red", color = "white", bold = TRUE)
```
##Avergae building book values 
```{r}
endowment_data <- read_rds(here("data", "endowment_filter_data_990.RDS"))
filed_schd_d <- endowment_data %>%
 mutate(Filled_Sch_D = ifelse(rowSums(across(CYBeginningYearBalanceAmt:TermEndowmentBalanceEOYPct), na.rm = TRUE) != 0, "Yes", "No")) %>% 
  select(EIN, Filled_Sch_D)

bookvalues_join <- bookvalues %>% 
  select(organization_name, EIN) %>%
  unique()

bookvalues_new <- bookvalues %>% 
  group_by(EIN) %>%
  summarise(mean = mean(BookValueSum)) %>% 
  left_join(bookvalues_join, by = "EIN") %>% 
  left_join(filed_schd_d, by = "EIN") %>% 
  filter(!is.na(Filled_Sch_D)) %>% 
  unique()

vis_averagebbv <- ggplot(bookvalues_new, aes(x = Filled_Sch_D, y = mean, color =as.factor(organization_name), fill =  as.factor(organization_name))) +
  geom_bar(position="dodge", stat='identity', alpha=0.3) +
  #theme_bw() +
  theme(legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7))+
  theme(plot.title = element_text(size = 13, face = "bold", hjust = .5),
        axis.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 10, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 10),
        strip.text = element_text(face="bold",size = 10)) +
  labs(y = "Average Schedule D Sum of Building Book Value",
       x = "With Or Without Enodowment",
       title = "Average Schedule D Sum of Building Book Value Change",
       color = "EIN") 

ggplotly(vis_averagebbv)

#Table 
bookvalues %>% 
  group_by(EIN) %>%
  summarise(mean = mean(BookValueSum)) %>% 
  left_join(bookvalues_join, by = "EIN") %>% 
  select(EIN, organization_name, mean) %>%
  arrange(desc(mean)) %>%
  kbl(caption = "Ranking of Average Building Book Values") %>%
  kable_material() %>%
  row_spec(row = 0, background = "red", color = "white", bold = TRUE)
```

## Regional Analysis
```{r}
test_state <- test_state %>% 
  filter(fiscal_year != "2023") %>%
  mutate(
    Time = substring(ReturnTs, 1, 10),
    TotalBookValueLandBuildingsAmt = as.numeric(TotalBookValueLandBuildingsAmt)
  )
```

```{r}
state_bookvalues <- ggplot(test_state, aes(x = factor(fiscal_year, level = c('2016','2017','2018','2019','2020','2021','2022')), y = as.numeric(test_state$TotalBookValueLandBuildingsAmt), color=as.factor(state)) )+
  geom_line(aes(group=1))+
  theme_bw() +
  theme(legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7))+
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 6),
        strip.text = element_text(face="bold",size = 5)) +
  labs(y = "Bookvalues of buildings and other lands",
       x = "Fiscal year") 
  
ggplotly(state_bookvalues)
```
```{r}
region_state <- read_csv(here("data", "region_state.csv")) %>% 
  select(state = "State Code", region = Region)
test_state_new<-test_state %>% 
  left_join(y = region_state, by = "state")
```

```{r}

state_bookvalues_new <- ggplot(test_state_new, aes(x = factor(fiscal_year, level = c('2016','2017','2018','2019','2020','2021','2022')), y = as.numeric(test_state$TotalBookValueLandBuildingsAmt),color=as.factor(state), fill=as.factor(state)))+
  geom_bar(position="dodge", stat='identity', alpha=0.3) +
  theme_bw() +
  theme(legend.key.size = unit(1, 'mm'),
        legend.text = element_text(size=7))+
  theme(plot.title = element_text(size = 10, face = "bold", hjust = .5),
        axis.title = element_text(size = 10, face = "bold"),
        plot.subtitle = element_text(size = 5, face = "italic", hjust = .5),
        axis.text.x = element_text(size = 10),
        axis.text.y = element_text(size = 6),
        strip.text = element_text(face="bold",size = 5)) +
  labs(y = "Bookvalues of buildings and other lands",
       x = "Fiscal year") + 
   facet_wrap(~region, scale="fixed") + 
  scale_x_discrete(breaks = c('2016','2017','2018','2019','2020','2021','2022'),labels=as.character(c("2016", "", "2018", "", "2020", "", "2022")))
  
ggplotly(state_bookvalues_new)
```


```{r}
# faceted graphs 
summary_state <- test_state %>% 
  group_by(state, fiscal_year, name) %>% 
  summarise(
    max = max(TotalBookValueLandBuildingsAmt),
    min = min(TotalBookValueLandBuildingsAmt),
    mean = mean(TotalBookValueLandBuildingsAmt),
    median = median(TotalBookValueLandBuildingsAmt),
    Number = n()
  )

ggplotly(
  summary_state %>% 
  ggplot(aes(x = fiscal_year, y = median, color = as.factor(state))) + 
    geom_line(aes(group=1))+ 
    theme_bw()+ 
    labs(title = "Median of Total Book Value of Land and Building each state")) 
```

```{r}
test_state %>% 
  group_by(state,fiscal_year) %>% 
  summarise(
    max = max(TotalBookValueLandBuildingsAmt, na.rm=TRUE),
    min = min(TotalBookValueLandBuildingsAmt, na.rm=TRUE),
    mean = mean(TotalBookValueLandBuildingsAmt, na.rm=TRUE),
    median = median(TotalBookValueLandBuildingsAmt, na.rm=TRUE),
  )
```


```{r}
ggplotly(
  test_state %>% filter(fiscal_year!=2014) %>% group_by(state, fiscal_year) %>% 
  summarise(
    max = max(TotalBookValueLandBuildingsAmt,na.rm = TRUE),
    min = min(TotalBookValueLandBuildingsAmt,na.rm = TRUE),
    mean = mean(TotalBookValueLandBuildingsAmt,na.rm = TRUE),
    median = median(TotalBookValueLandBuildingsAmt,na.rm = TRUE),
    Number = n()
  ) %>% 
  ggplot(aes(x = fiscal_year, y = max, color = factor(state))) + geom_point() + theme_bw()
)
```

```{r}
test_state %>% 
  group_by(state, fiscal_year) %>% 
  summarise(
    sum = sum(TotalBookValueLandBuildingsAmt)
  ) %>% 
  arrange(desc(sum))
```
