---
title: "Filtering out Filings"
author: "Rose Evard"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css', 'template.css')
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "output_html")})
---

```{css, echo = FALSE}

/* expand content to screenwidth rather than cutting it off */
#content{
max-width:2300px;
}

```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

library(tidyverse)
library(kableExtra)
library(here)
all_data <- readRDS(here("data", "data_990.RDS"))

```


```{r}

# make kable table with consistent formatting
make_table <- function(..., title = "") {
  title <- paste0("<center><span style = 'font-size:150%;color:black'><b>",
                  title,
                  "</span></b><center>")
   as_tibble(...) %>%
    kbl(caption = title) %>%
    kable_material() %>%
    row_spec(row=0, background = "#43494C" , color = "white", bold = TRUE)
}

```

## Filtering Amended 990s and 990T  

```{r, filtering}
## Retain only 990, 990 EZ's that are not marked as amended
filter_data <- all_data %>%
  filter(ReturnTypeCd == "990" | ReturnTypeCd == "990EZ")
```


## Companies with Multiple Filings in a Fiscal Year

There are `r filter_data %>% group_by(EIN, fiscal_year) %>% mutate(n=n()) %>% filter(n>1) %>% nrow()` companies that have multiple filings in a year. No company had more than 2 filings for a given fiscal year.

Additionally, in cases where there were 2 filings, one was always an amended filing, as we would expect.


```{r}

# check if one of the filings was amended for 
# every company that had multiple filings  

companies_mult_filings <- filter_data %>%
  group_by(EIN, fiscal_year) %>%
  mutate(n=n()) %>% 
  select(AmendedReturnInd, EIN, n, ReturnDate) %>%
  arrange(desc(n), EIN, fiscal_year)  %>% 
  filter(n>1) %>%
  mutate(amended_return_filled = ifelse(!is.na(AmendedReturnInd),1, 0)) %>%
  group_by(EIN, fiscal_year) %>%
  mutate(filled_out_ammended_total_for_year = sum(amended_return_filled)) %>%
  mutate(form_was_amended = case_when(
    filled_out_ammended_total_for_year == 1 ~ "Amended Once",
    filled_out_ammended_total_for_year == 0 ~ "Not Amended",
    filled_out_ammended_total_for_year > 1 ~ "Amended More than Once")) %>%
  arrange(desc(n), EIN, fiscal_year, AmendedReturnInd) %>%
  select(-c(amended_return_filled, filled_out_ammended_total_for_year))

```

```{r class.source = 'fold-show'}

companies_mult_filings %>%
  filter(form_was_amended != "Amended Once")
```

```{r}

# table showing companies with multiple filings per year
companies_mult_filings %>%
  select(`Fiscal Year` = fiscal_year,
         `Amended` = AmendedReturnInd,
         EIN, 
         `Form was Amended` = form_was_amended,
         `Return Date` = ReturnDate,
         `Number of Filings<br>for Fiscal Year` = n
         ) %>%
  make_table(title = "Companies with Multiple Filings in a Fiscal Year" ) %>%
  scroll_box(height = "400px")




```

```{r}

# notice we have 799 filings before filtering based on amended filings
# from the previous code there were 14 cases where a company filed 
# 2 times in a single fiscal year
filter_data %>% nrow() 

# take the amended version
filter_data_amended <- filter_data %>%
  mutate(amended_return_filled = ifelse(!is.na(AmendedReturnInd),1, 0)) %>%
  group_by(EIN, fiscal_year) %>%
  slice_max(order_by =amended_return_filled, n = 1)

# take the most recent return date
max_by_date <- filter_data %>%
  mutate(amended_return_filled = ifelse(!is.na(AmendedReturnInd),1, 0)) %>%
  group_by(EIN, fiscal_year) %>%
  slice_max(order_by =ReturnDate, n = 1)

# check that picking max return date is the same as picking the amended one
all.equal(filter_data_amended, max_by_date)


# now we have 799 - 14 = 785 observtions, as we would expect, 
# because we only took the 14 amended ones
nrow(filter_data_amended)

eins_with_multiple <- filter_data %>%
  group_by(EIN, fiscal_year) %>%
  mutate(n=n()) %>%
  filter(n > 1) %>%
  # need both fiscal year and ein number to check
  # which year/ein combinations had multiple filings
  mutate(EIN_year =paste0(EIN, fiscal_year)) %>%
  pull(EIN_year)
  

# of the eins with multiple filings, now we only have
# the ones that were ammended (AmendedReturnInd is "X", not NA)
filter_data_amended %>%
  mutate(EIN_year =paste0(EIN, fiscal_year)) %>%
  filter(EIN_year %in% eins_with_multiple) %>%
  filter(is.na(AmendedReturnInd)) 
  
```


## Detecting Groups who Changed Fiscal Year  

```{r identify bridging}
#Finding all companies with any short or abnormally long fiscal years  
#Note that there were not any long ones
short_years_filings <- filter_data_amended %>%
  select(EIN, TaxPeriodBeginDt, TaxPeriodEndDt) %>%
  mutate(year_len = difftime(TaxPeriodEndDt, TaxPeriodBeginDt, units = "days")) %>%
  filter(year_len < 364 | year_len > 365)

##Tabling it for the HTML
short_years_filings %>%
  kbl(caption = "Short fiscal years") %>%
  kable_material()
```

```{r,include=FALSE}

filter_data_amended %>%
  select(EIN, TaxPeriodBeginDt, TaxPeriodEndDt) %>%
  mutate(year_len = difftime(TaxPeriodEndDt,
                             TaxPeriodBeginDt, 
                             units = "days")) %>%
  ggplot(aes(x = year_len))+
  geom_histogram()

```



There is only one company which has a short fiscal year: 742371103

## Investigating single short fiscal year  

```{r}
#Filtering to 742371103, utilizing some of Quinn's code to graph gaps
filter_data_amended %>%
  filter(EIN == "742371103") %>%
  select(EIN, 
         BusinessName,
       TaxPeriodBeginDt, 
       TaxPeriodEndDt) %>%
  arrange(TaxPeriodBeginDt) %>%
  mutate(year_len = difftime(TaxPeriodEndDt, TaxPeriodBeginDt, units = "days"),
    before =  lag(TaxPeriodEndDt, n =1),
        gap = difftime(before, TaxPeriodBeginDt, units = "days"))%>%
  select(EIN, BusinessName, TaxPeriodBeginDt,TaxPeriodEndDt, year_len, gap) %>%
  kbl(caption = "Fiscal year length and fiscal gaps for EIN 742371103") %>%
  kable_material() %>%
  row_spec(row = 5, background = "red", color = "white", bold = TRUE)
```

As the shorter fiscal year is still a significant portion of the year, 302 days, I feel it should not be filtered.